---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/altair)](https://cran.r-project.org/package=altair)
[![Travis build status](https://travis-ci.org/ijlyttle/altair.svg?branch=master)](https://travis-ci.org/ijlyttle/altair)

# altair

The goal of altair is to help you build [Vega-Lite](https://vega.github.io/vega-lite) visualizations. Using the [reticulate](https://rstudio.github.io/reticulate) package, it provides an interface to the [Altair](https://altair-viz.github.io) Python library. This approach is different from efforts to build a native R interface to Vega-Lite, by Bob Rudis ([\@hrbrmstr](https://github.com/hrbrmstr)) and coworkers: the  [vegalite](https://github.com/hrbrmstr/vegalite) package, which has inspired this effort.

In this documentation, the capitalized word **Altair** shall refer to the Python API; the lower-case word **altair** shall refer to this R package.

## Installation

For this package, only a development version is available, from [GitHub](https://github.com/):

``` r
# install.packages("devtools")
devtools::install_github("ijlyttle/altair")
```
It is necessary to have a working Altair installation on your computer. As a first step, you may wish to  follow the [Altair installation instructions](https://altair-viz.github.io/getting_started/installation.html). Once you have Altair install in your Python environment, you can determine the path to your particular Python installation:

```python
# Python

import altair # make sure this works
import sys

sys.executable
```

It might return something like `/path/to/bin/python3.6`; you can use this value to set an environment variable to specify which Python environment to use. You might consider putting a line like this into your `.Renviron` file:

```
RETICULATE_PYTHON="/your/path/to/python3.6" # your path will be different
```

This is where the reticulate package will look first for Python on your computer.

Of course, once you have edited your `.Renviron` file, restart R for the changes to take effect.

## Example

This package provides a simple pass-through to the Altiar API. 

```{r example, eval=FALSE}
library("altair")

plot <- 
  alt$Chart(
    r_to_py(mtcars)
  )$mark_point(
  )$encode(
    x = "mpg:Q",
    y = "hp:Q",
    color = "cyl:N"
  )

vegalite(plot)
```

Some things to keep in mind:

  - Where you see a `.` in the Python examples, use a `$` instead.

  - Any data-frames you provide as arguments need to wrapped by `r_to_py()`.

  - In your data-frames, columns that contain dots, i.e. `Sepal.Width` will prevent Altair from compiling a chart spec.

Also, it mat remain to sort out how to get the `vegalite()` function to "do the right thing" when knitting to a non-html format. 

## Development plan

For the foreseeable future, this package is going to be very rough. At the moment, you are able to muck around with Vega-Lite 2.0. This means:

1. You can create charts by accessing the Python **Altair** API using **reticulate**.
1. You can display charts using the `vegalite()` htmlwidget.

There's really not much beyond that. Here's what I have in mind for the near future:

1. Tightening up existing capabilites.
1. A proper installation procedure for the Altair Python package, a function like `install_altair()`.

In the longer-term future it may be interesting to provide a proper R interface to the Python API; there are already some [encouraging first steps](https://github.com/ijlyttle/altair/issues/15) towards this. This package has a [manifesto](https://ijlyttle.github.io/altair/articles/manifesto.html) to outline some high-level ideas.

## Articles

The documentation for this package includes some articles:

- [First Example](https://ijlyttle.github.io/altair/articles/first-example.html): A walkthrough to get a first chart to work.

- [Interactive Examples](https://ijlyttle.github.io/altair/articles/interactive.html): A set of examples that work towards linked-brushing of two scatterplots.

- [Manifesto](https://ijlyttle.github.io/altair/articles/manifesto.html): A collection of ideas on where this package might go (and where it might want to avoid).

## Acknowledgements

This package does not contribute much new; it rests on these foundations:

- [Altair](https://altair-viz.github.io): Python interface to Vega-Lite
- [Reticulate](https://rstudio.github.io/reticulate): R framework to work with Python
- [Vega-Lite](https://vega.github.io/vega-lite): A grammar of interactive graphics
- [D3](https://d3js.org): Data-driven documents
- [vegalite](https://github.com/hrbrmstr/vegalite): Native R interface to Vega-Lite

A particular debt is owed to the [vegalite](https://github.com/hrbrmstr/vegalite) project as it provided a lot of the inspiration for this package.

## Code of conduct

Please note that this project is released with a [Contributor Code of Conduct](CODE_OF_CONDUCT.md).
By participating in this project you agree to abide by its terms.
