---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/altair)](https://cran.r-project.org/package=altair)
[![Travis build status](https://travis-ci.org/vegawidget/altair.svg?branch=master)](https://travis-ci.org/vegawidget/altair)

# altair

The goal of altair is to help you build [**Vega-Lite**](https://vega.github.io/vega-lite) visualizations. Using the [**reticulate**](https://rstudio.github.io/reticulate) package, it provides an interface to the [**Altair**](https://altair-viz.github.io) Python package. This approach is distinct from efforts to build a native R interface to Vega-Lite, by [Bob Rudis](https://github.com/hrbrmstr) and coworkers: the  [**vegalite**](https://github.com/hrbrmstr/vegalite) package, which has inspired this effort.

In this documentation, the capitalized word **Altair** shall refer to the Python package; the lower-case word **altair** shall refer to this R package.

## Installation

A development version is available from [GitHub](https://github.com/):

``` r
# install.packages("devtools")
devtools::install_github("vegawidget/altair") 
```

Because you are using a Python package, you may have some additional installation steps. These steps are described in greater detail in the [Installation](https://vegawidget.github.io/altair/articles/installation.html) article.

1. Python must be installed on your system. I recommend using a [Conda](https://conda.io/docs) installation; [Miniconda](https://conda.io/docs/user-guide/install/download.html#anaconda-or-miniconda) works well and installs more-quickly than Anaconda.

   If you work in a corporate or institutional environment, you may have to specify the location of your SSL certificate, or deal with a proxy. The installation article has a [section](https://vegawidget.github.io/altair/articles/installation.html#proxies) on this. 
   
2. Create a Conda environment called `"r-reticulate"`. The reticulate folks [recommend](https://rstudio.github.io/reticulate/articles/python_packages.html) standardizing on a common name for all packages that use reticulate. For more information, there is a [section](https://vegawidget.github.io/altair/articles/installation.html#python-env) in the installation article. 

3. Install Altair into your `"r-reticulate"` environment using `altair::install_altair()`. 

You may wish to add a line like this to the `.First()` function in your `.Rprofile`:

```r
reticulate::use_condaenv("r-reticulate")
```

The `use_condaenv()` function is called to provide a [hint to reticulate](https://rstudio.github.io/reticulate/articles/versions.html#order-of-discovery) on which Python environment to use.

## Example

```{r example}
library("altair")

vega_data <- import_vega_data()

chart <- 
  alt$Chart(vega_data$cars())$
  mark_point()$
  encode(
    x = "Horsepower:Q",
    y = "Miles_per_Gallon:Q",
    color = "Origin:N"
  )

chart
```

Some things to keep in mind:

  - Where you see a `.` in the Python examples, use a `$` instead.

  - In your data, columns that contain dots can be wrapped in square brackets in Altair specifications, e.g. `[Sepal.Width]`, to keep Altair from throwing an error.
  
These and other "gotchas" are compiled, along with workarounds, in an article: [Field Guide to Python Issues](https://vegawidget.github.io/altair/articles/field-guide-python.html).   

## Articles

The documentation for this package includes some [articles](https://vegawidget.github.io/altair/articles/index.html):

- [Installation](https://vegawidget.github.io/altair/articles/installation.html): some more-detailed instructions

- [Manifesto](https://vegawidget.github.io/altair/articles/manifesto.html): an outline of some high-level development ideas, which may change from time to time

Gallery:

- An adaptation to R of the [Altair Example Gallery](https://altair-viz.github.io/gallery), to demonstrate (not least to ourselves) that we are not missing any of the expressiveness of the Python API. You may be interested in the [Interactive Charts](https://vegawidget.github.io/altair/articles/example-gallery-08-interactive-charts.html) examples.  

Examples:

- [First Example](https://vegawidget.github.io/altair/articles/first-example.html): a walkthrough to get a first chart to work

- [Tooltips](https://vegawidget.github.io/altair/articles/tooltips.html): shows how Vega-Lite implements tooltips as an encoding within a chart, with formatting options

- [Vega Datasets](https://vegawidget.github.io/altair/articles/vega-datasets.html): work with [Vega datasets](https://github.com/altair-viz/vega_datasets) using `import_vega_data()`

- [View Composition](https://vegawidget.github.io/altair/articles/view-composition.html): how to facet, add layers to, repeat, and concatenate charts

- [Interactive Examples](https://vegawidget.github.io/altair/articles/interactive.html): a set of examples that work towards linked-brushing of two scatterplots

Field Guides:

- [Field Guide to Python Issues](https://vegawidget.github.io/altair/articles/field-guide-python.html): A place for "gotchas" and their workarounds

- [Field Guide to Rendering Charts](https://vegawidget.github.io/altair/articles/field-guide-rendering.html): How to use the `vegawidget()` function to render charts as HTML

## Acknowledgements

This package rests on these foundations:

- [Altair](https://altair-viz.github.io): Python interface to Vega-Lite

- [reticulate](https://rstudio.github.io/reticulate): R framework to work with Python

- [Vega-Lite](https://vega.github.io/vega-lite): a grammar of interactive graphics

- [htmlwidgets](https://www.htmlwidgets.org/): R framework to work with JavaScript visualizations

As well, a particular debt is owed to the folks behind the [vegalite](https://github.com/hrbrmstr/vegalite) package, as it provided a lot of the inspiration for these efforts.

This project is a collaborative effort. In addition to the principal author:

- [Alicia Schep](https://github.com/AliciaSchep) has contributed the concatenation functions, as well as sorted out many of the Python, JavaScript, and package-API issues.

- [Haley Jeppson](https://github.com/haleyjeppson) has adapted the [Altair Example Gallery](https://altair-viz.github.io/gallery) to R.

- [Heike Hofmann](https://github.com/heike) has been an invaluable advisor, providing incisive feedback, and insight into the fundamentals of interactive graphics.

## Contributing

Please note that this project is released with a [Contributor Code of Conduct](CODE_OF_CONDUCT.md).
By participating in this project you agree to abide by its terms. 

This project also has a [Contributing Guide](CONTRIBUTING.md).
