---
title: "Bar Charts"
author: "Haley Jeppson"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This document is adapted from the [Bar Charts](https://altair-viz.github.io/gallery/index.html#bar-charts) section of the Altair Example Gallery.

Our first step is to set up our environment:

```{r}
# devtools::install_github("vegawidget/altair")
library("altair")
library("tibble")
library("jsonlite")

vega_data <- import_vega_data()
```


## Bar Chart with Highlight

[Altair example](https://altair-viz.github.io/gallery/bar_chart_with_highlight.html)

#### Data

<details>
<summary>Definition</summary>
```{r}
data <- 
  data_frame(
    Day = 1:15, 
    Value = c(54.8, 112.1, 63.6, 37.6, 79.7, 137.9, 120.1, 103.3, 394.8,
              199.5, 72.3, 51.1, 112.0, 174.5, 130.5)
  )

data2 <- data_frame(ThresholdValue = 300, Threshold = "hazardous")
```
</details>

```{r}
glimpse(data)
```

```{r}
glimpse(data2)
```

#### Chart

```{r}
bar1 <- 
  alt$Chart(r_to_py(data))$
  mark_bar()$
  encode(
    x = "Day:O",
    y = "Value:Q"
  )

bar2 <- 
  alt$Chart(r_to_py(data))$
  mark_bar(color="#e45755")$
  encode(
    x = "Day:O",
    y = "baseline:Q",
    y2 = "Value:Q"
  )$
  transform_filter("datum.Value >= 300")$
  transform_calculate("baseline", "300")

rule <- 
  alt$Chart(r_to_py(data2))$
  mark_rule()$
  encode(y = "ThresholdValue:Q")

text <- 
  alt$Chart(r_to_py(data2))$
  mark_text(align="left", dx = 215, dy = -5)$
  encode(
    y = alt$Y("ThresholdValue:Q", axis = alt$Axis(title = "PM2.5 Value")),
    text = alt$value("hazardous")
  )

chart <- (bar1 + text + bar2 + rule)

chart
```

## Box Plot with Min/Max Whiskers

[Altair example](https://altair-viz.github.io/gallery/boxplot_max_min.html)

#### Data

```{r}
glimpse(vega_data$population())
```

#### Chart

```{r}
# Define aggregate fields
lower_box <- "q1(people):Q"
lower_whisker <- "min(people):Q"
upper_box <- "q3(people):Q"
upper_whisker <- "max(people):Q"

# Compose each layer individually
lower_plot <- 
  alt$Chart(r_to_py(vega_data$population()))$
  mark_rule()$
  encode(
    y = alt$Y(lower_whisker, axis = alt$Axis(title = "population")),
    y2 = lower_box,
    x = "age:O"
  )

middle_plot <- 
  alt$Chart(r_to_py(vega_data$population()))$
  mark_bar(size = 5.0)$
  encode(
    y = lower_box,
    y2 = upper_box,
    x ="age:O"
  )

upper_plot <- 
  alt$Chart(r_to_py(vega_data$population()))$
  mark_rule()$
  encode(
    y = upper_whisker,
    y2 = upper_box,
    x = "age:O"
  )

middle_tick <- 
  alt$Chart(r_to_py(vega_data$population()))$
  mark_tick(color = "white", size = 5.0)$
  encode(
    y = "median(people):Q",
    x = "age:O"
  )

chart <- (lower_plot + middle_plot + upper_plot + middle_tick)

chart
```

## Calculating Percentage of Total

[Altair example](https://altair-viz.github.io/gallery/percentage_of_total.html)

TODO: put a note here about the `**` in the Python code. 

#### Data

<details>
<summary>Definition</summary>
```{r}
activities <- 
  data_frame(
    Activity = c("Sleeping", "Eating", "TV", "Work", "Exercise"),
    Time = c(8, 2, 4, 8, 2)
  )
```
</details>

```{r}
glimpse(activities)
```

#### Chart

```{r}
chart <- 
  alt$Chart(r_to_py(activities))$
  mark_bar()$
  encode(
    alt$X("PercentOfTotal:Q", axis = alt$Axis(format = ".0%")),
    y = "Activity:N"
  )$
  transform_window(
    window = list(alt$WindowFieldDef(op = "sum", field = "Time", as = "TotalTime")),
    frame = list(NULL, NULL)
  )$
  transform_calculate(PercentOfTotal = "datum.Time / datum.TotalTime")

chart
```

## Candlestick Chart

[Altair example](https://altair-viz.github.io/gallery/candlestick_chart.html)

#### Data

<details>
<summary>Definition</summary>
```{r}
data <- 
jsonlite::fromJSON('[
  {
    "date": "01-Jun-2009",
    "open": 28.7,
    "high": 30.05,
    "low": 28.45,
    "close": 30.04,
    "signal": "short",
    "ret": -4.89396411092985
  },
  {
    "date": "02-Jun-2009",
    "open": 30.04,
    "high": 30.13,
    "low": 28.3,
    "close": 29.63,
    "signal": "short",
    "ret": -0.322580645161295
  },
  {
    "date": "03-Jun-2009",
    "open": 29.62,
    "high": 31.79,
    "low": 29.62,
    "close": 31.02,
    "signal": "short",
    "ret": 3.68663594470045
  },
  {
    "date": "04-Jun-2009",
    "open": 31.02,
    "high": 31.02,
    "low": 29.92,
    "close": 30.18,
    "signal": "short",
    "ret": 4.51010886469673
  },
  {
    "date": "05-Jun-2009",
    "open": 29.39,
    "high": 30.81,
    "low": 28.85,
    "close": 29.62,
    "signal": "short",
    "ret": 6.08424336973478
  },
  {
    "date": "08-Jun-2009",
    "open": 30.84,
    "high": 31.82,
    "low": 26.41,
    "close": 29.77,
    "signal": "short",
    "ret": 1.2539184952978
  },
  {
    "date": "09-Jun-2009",
    "open": 29.77,
    "high": 29.77,
    "low": 27.79,
    "close": 28.27,
    "signal": "short",
    "ret": -5.02431118314424
  },
  {
    "date": "10-Jun-2009",
    "open": 26.9,
    "high": 29.74,
    "low": 26.9,
    "close": 28.46,
    "signal": "short",
    "ret": -5.46623794212217
  },
  {
    "date": "11-Jun-2009",
    "open": 27.36,
    "high": 28.11,
    "low": 26.81,
    "close": 28.11,
    "signal": "short",
    "ret": -8.3743842364532
  },
  {
    "date": "12-Jun-2009",
    "open": 28.08,
    "high": 28.5,
    "low": 27.73,
    "close": 28.15,
    "signal": "short",
    "ret": -5.52763819095477
  },
  {
    "date": "15-Jun-2009",
    "open": 29.7,
    "high": 31.09,
    "low": 29.64,
    "close": 30.81,
    "signal": "long",
    "ret": 3.4920634920635
  },
  {
    "date": "16-Jun-2009",
    "open": 30.81,
    "high": 32.75,
    "low": 30.07,
    "close": 32.68,
    "signal": "short",
    "ret": 0.155038759689914
  },
  {
    "date": "17-Jun-2009",
    "open": 31.19,
    "high": 32.77,
    "low": 30.64,
    "close": 31.54,
    "signal": "short",
    "ret": 5.82822085889571
  },
  {
    "date": "18-Jun-2009",
    "open": 31.54,
    "high": 31.54,
    "low": 29.6,
    "close": 30.03,
    "signal": "short",
    "ret": 8.17610062893082
  },
  {
    "date": "19-Jun-2009",
    "open": 29.16,
    "high": 29.32,
    "low": 27.56,
    "close": 27.99,
    "signal": "short",
    "ret": 8.59872611464968
  },
  {
    "date": "22-Jun-2009",
    "open": 30.4,
    "high": 32.05,
    "low": 30.3,
    "close": 31.17,
    "signal": "short",
    "ret": 15.4907975460123
  },
  {
    "date": "23-Jun-2009",
    "open": 31.3,
    "high": 31.54,
    "low": 27.83,
    "close": 30.58,
    "signal": "short",
    "ret": 11.7370892018779
  },
  {
    "date": "24-Jun-2009",
    "open": 30.58,
    "high": 30.58,
    "low": 28.79,
    "close": 29.05,
    "signal": "long",
    "ret": -10.4234527687296
  },
  {
    "date": "25-Jun-2009",
    "open": 29.45,
    "high": 29.56,
    "low": 26.3,
    "close": 26.36,
    "signal": "long",
    "ret": 0
  },
  {
    "date": "26-Jun-2009",
    "open": 27.09,
    "high": 27.22,
    "low": 25.76,
    "close": 25.93,
    "signal": "long",
    "ret": 0
  },
  {
    "date": "29-Jun-2009",
    "open": 25.93,
    "high": 27.18,
    "low": 25.29,
    "close": 25.35,
    "signal": "long",
    "ret": 5.26315789473684
  },
  {
    "date": "30-Jun-2009",
    "open": 25.36,
    "high": 27.38,
    "low": 25.02,
    "close": 26.35,
    "signal": "long",
    "ret": 6.73758865248228
  }
]')
```
</details>

```{r}
glimpse(data)
```

#### Chart

```{r}
open_close_color <- 
  alt$condition(
   "datum.open < datum.close",
   alt$value("#06982d"),
   alt$value("#ae1325")
  )

rule <- 
  alt$Chart(r_to_py(data))$
  mark_rule()$
  encode(
    alt$X(
      "date:T",
      timeUnit = "yearmonthdate",
      scale = alt$Scale(
        domain = list(
          list(month= 5, date= 31, year= 2009),
          list(month= 7, date= 1, year= 2009)
        )
      ),
      axis = alt$Axis(format="%m/%d", title="Date in 2009")
    ),
    alt$Y(
      "low",
      scale = alt$Scale(zero=FALSE),
      axis = alt$Axis(title="Price")
    ),
    alt$Y2("high"),
    color = open_close_color
)

bar <- 
  alt$Chart(r_to_py(data))$
  mark_bar()$
  encode(
    alt$X("date:T", timeUnit = "yearmonthdate"),
    y = "open",
    y2 = "close",
    color = open_close_color
  )

chart <- (rule + bar)

chart
```


## Diverging Stacked Bar Chart

[Altair example](https://altair-viz.github.io/gallery/diverging_stacked_bar_chart.html)

#### Data

<details>
<summary>Definition</summary>
```{r}
data <- fromJSON('[
  {
    "question": "Question 1",
    "type": "Strongly disagree",
    "value": 24,
    "percentage": 0.7,
    "percentage_start": -19.1,
    "percentage_end": -18.4
  },
  {
    "question": "Question 1",
    "type": "Disagree",
    "value": 294,
    "percentage": 9.1,
    "percentage_start": -18.4,
    "percentage_end": -9.2
  },
  {
    "question": "Question 1",
    "type": "Neither agree nor disagree",
    "value": 594,
    "percentage": 18.5,
    "percentage_start": -9.2,
    "percentage_end": 9.2
  },
  {
    "question": "Question 1",
    "type": "Agree",
    "value": 1927,
    "percentage": 59.9,
    "percentage_start": 9.2,
    "percentage_end": 69.2
  },
  {
    "question": "Question 1",
    "type": "Strongly agree",
    "value": 376,
    "percentage": 11.7,
    "percentage_start": 69.2,
    "percentage_end": 80.9
  },
  
  {
    "question": "Question 2",
    "type": "Strongly disagree",
    "value": 2,
    "percentage": 18.2,
    "percentage_start": -36.4,
    "percentage_end": -18.2
  },
  {
    "question": "Question 2",
    "type": "Disagree",
    "value": 2,
    "percentage": 18.2,
    "percentage_start": -18.2,
    "percentage_end": 0
  },
  {
    "question": "Question 2",
    "type": "Neither agree nor disagree",
    "value": 0,
    "percentage": 0,
    "percentage_start": 0,
    "percentage_end": 0
  },
  {
    "question": "Question 2",
    "type": "Agree",
    "value": 7,
    "percentage": 63.6,
    "percentage_start": 0,
    "percentage_end": 63.6
  },
  {
    "question": "Question 2",
    "type": "Strongly agree",
    "value": 11,
    "percentage": 0,
    "percentage_start": 63.6,
    "percentage_end": 63.6
  },
  
  {
    "question": "Question 3",
    "type": "Strongly disagree",
    "value": 2,
    "percentage": 20,
    "percentage_start": -30,
    "percentage_end": -10
  },
  {
    "question": "Question 3",
    "type": "Disagree",
    "value": 0,
    "percentage": 0,
    "percentage_start": -10,
    "percentage_end": -10
  },
  {
    "question": "Question 3",
    "type": "Neither agree nor disagree",
    "value": 2,
    "percentage": 20,
    "percentage_start": -10,
    "percentage_end": 10
  },
  {
    "question": "Question 3",
    "type": "Agree",
    "value": 4,
    "percentage": 40,
    "percentage_start": 10,
    "percentage_end": 50
  },
  {
    "question": "Question 3",
    "type": "Strongly agree",
    "value": 2,
    "percentage": 20,
    "percentage_start": 50,
    "percentage_end": 70
  },
  
  {
    "question": "Question 4",
    "type": "Strongly disagree",
    "value": 0,
    "percentage": 0,
    "percentage_start": -15.6,
    "percentage_end": -15.6
  },
  {
    "question": "Question 4",
    "type": "Disagree",
    "value": 2,
    "percentage": 12.5,
    "percentage_start": -15.6,
    "percentage_end": -3.1
  },
  {
    "question": "Question 4",
    "type": "Neither agree nor disagree",
    "value": 1,
    "percentage": 6.3,
    "percentage_start": -3.1,
    "percentage_end": 3.1
  },
  {
    "question": "Question 4",
    "type": "Agree",
    "value": 7,
    "percentage": 43.8,
    "percentage_start": 3.1,
    "percentage_end": 46.9
  },
  {
    "question": "Question 4",
    "type": "Strongly agree",
    "value": 6,
    "percentage": 37.5,
    "percentage_start": 46.9,
    "percentage_end": 84.4
  },
  
  {
    "question": "Question 5",
    "type": "Strongly disagree",
    "value": 0,
    "percentage": 0,
    "percentage_start": -10.4,
    "percentage_end": -10.4
  },
  {
    "question": "Question 5",
    "type": "Disagree",
    "value": 1,
    "percentage": 4.2,
    "percentage_start": -10.4,
    "percentage_end": -6.3
  },
  {
    "question": "Question 5",
    "type": "Neither agree nor disagree",
    "value": 3,
    "percentage": 12.5,
    "percentage_start": -6.3,
    "percentage_end": 6.3
  },
  {
    "question": "Question 5",
    "type": "Agree",
    "value": 16,
    "percentage": 66.7,
    "percentage_start": 6.3,
    "percentage_end": 72.9
  },
  {
    "question": "Question 5",
    "type": "Strongly agree",
    "value": 4,
    "percentage": 16.7,
    "percentage_start": 72.9,
    "percentage_end": 89.6
  },
  
  {
    "question": "Question 6",
    "type": "Strongly disagree",
    "value": 1,
    "percentage": 6.3,
    "percentage_start": -18.8,
    "percentage_end": -12.5
  },
  {
    "question": "Question 6",
    "type": "Disagree",
    "value": 1,
    "percentage": 6.3,
    "percentage_start": -12.5,
    "percentage_end": -6.3
  },
  {
    "question": "Question 6",
    "type": "Neither agree nor disagree",
    "value": 2,
    "percentage": 12.5,
    "percentage_start": -6.3,
    "percentage_end": 6.3
  },
  {
    "question": "Question 6",
    "type": "Agree",
    "value": 9,
    "percentage": 56.3,
    "percentage_start": 6.3,
    "percentage_end": 62.5
  },
  {
    "question": "Question 6",
    "type": "Strongly agree",
    "value": 3,
    "percentage": 18.8,
    "percentage_start": 62.5,
    "percentage_end": 81.3
  },
  
  {
    "question": "Question 7",
    "type": "Strongly disagree",
    "value": 0,
    "percentage": 0,
    "percentage_start": -10,
    "percentage_end": -10
  },
  {
    "question": "Question 7",
    "type": "Disagree",
    "value": 0,
    "percentage": 0,
    "percentage_start": -10,
    "percentage_end": -10
  },
  {
    "question": "Question 7",
    "type": "Neither agree nor disagree",
    "value": 1,
    "percentage": 20,
    "percentage_start": -10,
    "percentage_end": 10
  },
  {
    "question": "Question 7",
    "type": "Agree",
    "value": 4,
    "percentage": 80,
    "percentage_start": 10,
    "percentage_end": 90
  },
  {
    "question": "Question 7",
    "type": "Strongly agree",
    "value": 0,
    "percentage": 0,
    "percentage_start": 90,
    "percentage_end": 90
  },
  
  {
    "question": "Question 8",
    "type": "Strongly disagree",
    "value": 0,
    "percentage": 0,
    "percentage_start": 0,
    "percentage_end": 0
  },
  {
    "question": "Question 8",
    "type": "Disagree",
    "value": 0,
    "percentage": 0,
    "percentage_start": 0,
    "percentage_end": 0
  },
  {
    "question": "Question 8",
    "type": "Neither agree nor disagree",
    "value": 0,
    "percentage": 0,
    "percentage_start": 0,
    "percentage_end": 0
  },
  {
    "question": "Question 8",
    "type": "Agree",
    "value": 0,
    "percentage": 0,
    "percentage_start": 0,
    "percentage_end": 0
  },
  {
    "question": "Question 8",
    "type": "Strongly agree",
    "value": 2,
    "percentage": 100,
    "percentage_start": 0,
    "percentage_end": 100
  }
]')
```
</details>

```{r}
glimpse(data)
```

#### Chart

```{r}
color_scale <-
  alt$Scale(
    domain = list(
      "Strongly disagree",
      "Disagree",
      "Neither agree nor disagree",
      "Agree",
      "Strongly agree"
    ),
    range = list("#c30d24", "#f3a583", "#cccccc", "#94c6da", "#1770ab")
  )

y_axis <-
  alt$Axis(
    title = "Question",
    offset = 5,
    ticks = FALSE,
    minExtent = 60,
    domain = FALSE
  )

chart <- 
  alt$Chart(r_to_py(data))$
  mark_bar()$
  encode(
    x = "percentage_start:Q",
    x2 = "percentage_end:Q",
    y = alt$Y("question:N", axis = y_axis),
    color = alt$Color(
      "type:N",
      legend = alt$Legend(title = "Response"),
      scale = color_scale
    )
  )

chart
```

## Error Bars showing Confidence Interval

[Altair example](https://altair-viz.github.io/gallery/error_bars_with_ci.html)

#### Data

```{r}
glimpse(vega_data$barley())
```

#### Chart

```{r}
points <- 
  alt$Chart(r_to_py(vega_data$barley()))$
  mark_point(filled = TRUE)$
  encode(
    x = alt$X(
      "mean(yield)",
      scale = alt$Scale(zero = FALSE),
      axis = alt$Axis(title = "Barley Yield")
    ),
    y = "variety",
    color = alt$value("black")
  )

error_bars <- 
  alt$Chart(r_to_py(vega_data$barley()))$
  mark_rule()$
  encode(
    x = "ci0(yield)",
    x2 = "ci1(yield)",
    y = "variety"
  )

chart <- (points + error_bars)

chart
```

## Gantt Chart

[Altair example](https://altair-viz.github.io/gallery/gantt_chart.html)

#### Data

<details>
<summary>Definition</summary>
```{r}
data <- 
  fromJSON('[
    {"task": "A", "start": 1, "end": 3},
    {"task": "B", "start": 3, "end": 8},
    {"task": "C", "start": 8, "end": 10}
  ]')
```
</details>

```{r}
glimpse(data)
```

#### Chart

```{r}
chart <- 
  alt$Chart(r_to_py(data))$
  mark_bar()$
  encode(
    x = "start",
    x2 = "end",
    y = "task"
  )

chart
```

## Grouped Bar Chart

[Altair example](https://altair-viz.github.io/gallery/grouped_bar_chart.html)

TODO: It may be useful to note which of these are JavaScript expressions, and to come up with a convention to denote that. For example, what is the value of wrapping in `JS()`? Is it just to denote this to ourselves? This may be reason enough to do this. Also try to figure out what is up with the `if_()` notation in Python. 


```{r}
chart <- 
  alt$Chart(r_to_py(vega_data$population()))$
  mark_bar(stroke = "transparent")$
  encode(
    x = alt$X(
      "gender:N", 
      scale = alt$Scale(rangeStep = 12), 
      axis = alt$Axis(title = "")
    ),
    y = alt$Y(
      "sum(people):Q", 
      axis=alt$Axis(title = "population", grid = FALSE)
    ),
    color = alt$Color(
      "gender:N", 
      scale = alt$Scale(range = list("#EA98D2", "#659CCA"))
    ),
    column="age:O"
  )$
  configure_view(stroke = "transparent")$
  configure_axis(domainWidth = 0.8)$
  transform_filter(JS("datum.year == 2000"))$
  transform_calculate("gender", JS('if(datum.sex == 2, "Female", "Male")'))$
  copy()

chart
```

## Horizontal Aggregate Bar Chart

[Altair example](https://altair-viz.github.io/gallery/aggregate_bar_chart.html)

Need to specify the `transform_filter` as a string of JavaScript

#### Data

```{r}
glimpse(vega_data$population())
```

#### Chart

```{r}
chart <- 
  alt$Chart(r_to_py(vega_data$population()))$
  mark_bar()$
  encode(
    x = alt$X("sum(people):Q", axis = alt$Axis(title = "population")),
    y = "age:O"
  )$
  properties(height = 300, width = 300)$
  transform_filter(JS("datum.year == 2000"))

chart
```

## Horizontal Stacked Bar Chart

[Altair example](https://altair-viz.github.io/gallery/horizontal_stacked_bar_chart.html)

#### Data

```{r}
glimpse(vega_data$barley())
```

#### Chart

```{r}
chart <- 
  alt$Chart(r_to_py(vega_data$barley()))$
  mark_bar()$
  encode(
    x = "sum(yield)",
    y = "variety",
    color = "site"
  )

chart
```

## LayerChart with Bar and Tick

[Altair example](https://altair-viz.github.io/gallery/layered_chart_bar_mark.html)

#### Data

<details>
<summary>Definition</summary>
```{r}
data <- 
  data_frame(
    project = c("a", "b", "c", "d", "e", "f", "g"),
    score = c(25, 57, 23, 19, 8, 47, 8),
    goal = c(25, 47, 30, 27, 38, 19, 4)
  )
```
</details>

```{r}
glimpse(data)
```

#### Chart

```{r}
a <- 
  alt$Chart()$
  mark_bar()$
  encode(
    x = "project",
    y = "score"
  )

b <- 
  alt$Chart()$
  mark_tick(color = "red")$
  encode(
    x = "project",
    y = "goal"
  )

chart <- 
  (a + b)$ 
  properties(data = r_to_py(data))$
  # controls the width of the tick
  configure_tick(thickness = 2, bandSize=35)$
  # controls the width of the bar
  configure_scale(rangeStep = 40)

chart
```

## Layered Bar Chart

[Altair example](https://altair-viz.github.io/gallery/layered_bar_chart.html)

#### Data

```{r}
glimpse(vega_data$population())
```

#### Chart

```{r}
chart <- 
  alt$Chart(r_to_py(vega_data$population()))$
  mark_bar(opacity = 0.7)$
  encode(
    x = alt$X("age:O", scale = alt$Scale(rangeStep = 17)),
    y = alt$Y(
      "sum(people):Q", 
      axis = alt$Axis(title = "population"), 
      stack = NULL
    ),
    color = alt$Color(
      "gender:N", 
      scale = alt$Scale(range = list("#EA98D2", "#659CCA")))
  )$
  transform_filter(JS("datum.year == 2000"))$
  transform_calculate("gender", JS('if(datum.sex == 2, "Female", "Male")'))

chart
```

## Layered Bar Chart

[Altair example](https://altair-viz.github.io/gallery/layered_bar_chart.html)

## Layered Bar Chart with Line as Mean

[Altair example](https://altair-viz.github.io/gallery/mean_overlay_over_precipitiation_chart.html)

#### Data

```{r}
glimpse(vega_data$seattle_weather())
```

#### Chart

```{r}
source <- r_to_py(vega_data$seattle_weather())

bar <- 
  alt$Chart(source)$
  mark_bar()$
  encode(
    x = alt$X("date:O", timeUnit = "month"),
    y = alt$Y("mean(precipitation):Q")
  )

rule <- 
  alt$Chart(source)$
  mark_rule(color = "red")$
  encode(
    y = "mean(precipitation)",
    size = alt$value(3)
  )

chart <- (bar + rule)

chart
```

## Layered Plot with Dual-Axis

[Altair example](https://altair-viz.github.io/gallery/layered_plot_with_dual_axis.html) 

TODO: find a diplomatic way to say that this seems to support Hadley's (and others') edict on dual axes.

#### Data

```{r}
glimpse(vega_data$seattle_weather())
```

#### Chart

```{r}
base <- 
  alt$Chart(r_to_py(vega_data$seattle_weather()))$
  encode(
    x = alt$X(
      "date:O",
      axis = alt$Axis(format = "%b"),
      timeUnit = "month",
      scale = alt$Scale(zero = FALSE)
    )
  )

bar <- 
  base$
  mark_bar()$
  encode(
    y = "mean(precipitation)"
  )


line <-  
  base$
  mark_line(color = "red")$
  encode(
    y="mean(temp_max)"
  )

chart <-
  (bar + line)$
  resolve_scale(y= "independent")

chart
```

## Normalized Stacked Bar Chart

[Altair example](https://altair-viz.github.io/gallery/normalized_stacked_bar_chart.html)

#### Data

```{r}
glimpse(vega_data$population())
```

#### Chart

```{r}
chart <- 
  alt$Chart(r_to_py(vega_data$population()))$
  mark_bar()$
  encode(
    x = alt$X("age:O", scale = alt$Scale(rangeStep = 17)),
    y = alt$Y(
      "sum(people):Q",
       axis = alt$Axis(title="population"),
       stack = "normalize"
    ),
    color = alt$Color(
      "gender:N",
      scale = alt$Scale(range = c("#EA98D2", "#659CCA"))
    )
  )$
  transform_filter(JS("datum.year == 2000"))$
  transform_calculate("gender", JS('if(datum.sex == 2, "Female", "Male")'))

chart
```

## Simple Bar Chart with Labels

[Altair example](https://altair-viz.github.io/gallery/bar_chart_with_labels.html)

#### Data

<details>
<summary>Definition</summary>
```{r}
data <- 
  data_frame(
    a = c("A", "B", "C", "D", "E", "F", "G", "H", "I"),
    b = c(28, 55, 43, 91, 81, 53, 19, 87, 52)
  )
```
</details>

```{r}
glimpse(data)
```

#### Chart

```{r}
bars <- 
  alt$Chart(r_to_py(data))$
  mark_bar()$
  encode(
    x = "b",
    y = "a"
  )

text <- 
  bars$
  mark_text(align = "left", baseline = "middle", dx = 3)$
  encode(
    text = "b"
  )

chart <- (bars + text)

chart
```

## Stacked Bar Chart

[Altair example](https://altair-viz.github.io/gallery/stacked_bar_chart.html)


```{r}
glimpse(vega_data$seattle_weather())
```

#### Chart
```{r}
chart <- 
  alt$Chart(r_to_py(vega_data$seattle_weather()))$
  mark_bar()$
  encode(
    x = alt$X(
      "date:N",
      axis = alt$Axis(title = "Month of the Year"),
      timeUnit = "month"
    ),
    y = "count(weather)",
    color = alt$Color(
      "weather:N",
      legend = alt$Legend(title = "Weather type"),
      scale = alt$Scale(
        domain = list("sun", "fog", "drizzle", "rain", "snow"),
        range = list("#e7ba42", "#c7c7c7", "#aec7e8", "#1f77b4", "#9467bd")
      )
    )
  )

chart
```

## Trellis Stacked Bar Chart

[Altair example](https://altair-viz.github.io/gallery/trellis_stacked_bar_chart.html)

```{r}
glimpse(vega_data$barley())
```

#### Chart
```{r}
chart <- 
  alt$Chart(r_to_py(vega_data$barley()))$
  mark_bar()$
  encode(
    x = "sum(yield)",
    y = "variety",
    color = "site",
    column = "year"
  )$
  properties(width = 250)

chart
```












