---
title: "Scatter Plots"
author: "Haley Jeppson"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This document is adapted from the [Scatter Plots](https://altair-viz.github.io/gallery/index.html#scatter-plots) section of the Altair Example Gallery.

Our first step is to set up our environment:

```{r message=FALSE}
library("altair")
library("tibble")
library("dplyr")
library("purrr")
library("tidyr")
library("jsonlite")

vega_data <- import_vega_data()
```


## Binned Scatterplot

[Altair example](https://altair-viz.github.io/gallery/binned_scatterplot.html)

#### Data

```{r}
glimpse(fromJSON(vega_data$movies$url))
```

#### Chart

```{r}
chart <- 
  alt$Chart(vega_data$movies$url)$
  mark_circle()$
  encode(
    alt$X("IMDB_Rating:Q", bin = TRUE),
    alt$Y("Rotten_Tomatoes_Rating:Q", bin = TRUE),
    size = "count()"
  )

chart
```

## Brushing Scatter Plot to show data on a table


[Altair example](https://altair-viz.github.io/gallery/scatter_linked_table.html)

#### Data

```{r}
glimpse(vega_data$cars())
```

#### Chart

```{r}
source <- vega_data$cars()

# Brush for selection
brush <- alt$selection(type = "interval")

# Scatter Plot
points <- 
  alt$Chart(source)$
  mark_point()$
  encode(
    x = "Horsepower:Q",
    y = "Miles_per_Gallon:Q",
    color = alt$condition(brush, "Cylinders:O", alt$value("grey"))
  )$add_selection(brush)

# Base chart for data tables
ranked_text <-
  alt$Chart(source)$
  mark_text()$
  encode(
    y = alt$Y('row_number:O', axis = NULL)
  )$
  transform_window(row_number = "row_number()")$
  transform_filter(brush)$
  transform_window(rank = "rank(row_number)")$
  transform_filter("datum.rank<20")

# Data Tables
horsepower <- 
  ranked_text$
  encode(
    text = "Horsepower:N"
  )$
  properties(title = "Horsepower")

mpg <- 
  ranked_text$
  encode(
    text = "Miles_per_Gallon:N"
  )$
  properties(title = "MPG")

origin <- 
  ranked_text$
  encode(
    text = "Origin:N"
  )$
  properties(title = "Origin")

text <- (horsepower | mpg | origin) # Combine data tables

# Build chart
chart <- (points | text)$resolve_legend(color = "independent")

chart
```


## Bubble Plot

[Altair example](https://altair-viz.github.io/gallery/bubble_plot.html)

#### Data

```{r}
glimpse(vega_data$cars())
```

#### Chart

```{r}
chart <- 
  alt$Chart(vega_data$cars())$
  mark_point()$
  encode(
    x = "Horsepower:Q",
    y = "Miles_per_Gallon:Q",
    size = "Acceleration"
  )

chart
```

## Connected Scatterplot (Lines with Custom Paths)

[Altair example](https://altair-viz.github.io/gallery/connected_scatterplot.html)

> This example shows how layering can be used to build a plot. This dataset tracks miles driven per capita along with gas prices annually from 1956 to 2010. It is based on the May 2, 2010 New York Times article ‘Driving Shifts Into Reverse’. See this [reference](http://mbostock.github.io/protovis/ex/driving.html).

#### Data

```{r}
glimpse(vega_data$driving())
```

#### Chart

```{r}
lines <- 
  alt$Chart(vega_data$driving())$
  mark_line()$
  encode(
    x = alt$X("miles", scale = alt$Scale(zero = FALSE)),
    y = alt$Y("gas", scale = alt$Scale(zero = FALSE)),
    order = "year"
  )

points <- 
  alt$Chart(vega_data$driving())$
  mark_circle()$
  encode(
    alt$X("miles", scale = alt$Scale(zero = FALSE)),
    alt$Y("gas", scale = alt$Scale(zero = FALSE))
  )

chart <- (lines + points)

chart
```

## Dot Dash Plot

[Altair example](https://altair-viz.github.io/gallery/dot_dash_plot.html)

> This example shows how to make a dot-dash plot presented in Edward Tufte’s book Visual Display of Quantitative Information on page 133. This example is based on [g3o2's block](https://bl.ocks.org/g3o2/bd4362574137061c243a2994ba648fb8).

#### Data

```{r}
glimpse(vega_data$cars())
```

#### Chart

```{r}
cars <- vega_data$cars()

brush <- alt$selection(type = "interval")

tick_axis <- alt$Axis(labels=FALSE, domain=FALSE, ticks=FALSE)

tick_axis_notitle <- 
  alt$Axis(labels=FALSE, domain=FALSE, ticks=FALSE, title="")

points <-
  alt$Chart(cars)$
  mark_point()$
  encode(
    x = alt$X("Miles_per_Gallon", axis = alt$Axis(title = "")),
    y = alt$Y("Horsepower", axis = alt$Axis(title = "")),
    color = alt$condition(brush, "Origin", alt$value("grey"))
  )$
  properties(selection = brush)

x_ticks <- 
  alt$Chart(cars)$
  mark_tick()$
  encode(
    x = alt$X("Miles_per_Gallon", axis = tick_axis),
    y = alt$Y("Origin", axis = tick_axis_notitle),
    color = alt$condition(brush, "Origin", alt$value("lightgrey"))
  )$
  properties(selection = brush)

y_ticks <- 
  alt$Chart(cars)$
  mark_tick()$
  encode(
    alt$X("Origin", axis = tick_axis_notitle),
    alt$Y("Horsepower", axis = tick_axis),
    color=alt$condition(brush, "Origin", alt$value("lightgrey"))
  )$
  properties(selection = brush)

chart <- (y_ticks | (points & x_ticks))

chart
```

## Multifeature Scatter Plot

[Altair example](https://altair-viz.github.io/gallery/multifeature_scatter_plot.html)

#### Data

We will use the cars dataset here.

```{r}
glimpse(vega_data$cars())
```

#### Chart

```{r}
chart <- 
  alt$Chart(vega_data$cars())$
  mark_circle()$
  encode(
    x = alt$X("Weight_in_lbs", scale = alt$Scale(zero = FALSE)),
    y = alt$Y("Miles_per_Gallon", scale = alt$Scale(zero = FALSE, padding = 1)),
    color = alt$Color("Cylinders:N"),
    size = "Displacement"
  )

chart
```



## Polynomial Fit Plot

[Altair example](https://altair-viz.github.io/gallery/poly_fit.html)

#### Data

<details>
<summary>Definition</summary>
```{r}
n <- 40

data <- 
  tibble(
    x = runif(n)**2,
    y = 10 - 1 / (x + 0.1) + runif(n)
  )

degree <- c(1, 3, 5)

model_poly <- 
  map(degree, ~lm(y ~ poly(x, degree = .x), data = data)) 

x_pred <- seq(min(data$x), max(data$x), length.out = 500)

data_predict <- 
  crossing(degree, x = x_pred) %>%
  nest(x) %>%
  mutate(y = map2(model_poly, data, predict)) %>%
  unnest()
```
</details>


```{r}
glimpse(data)
```

```{r}
glimpse(data_predict)
```

#### Chart

```{r}
chart_data <- 
  alt$Chart(data)$
  mark_circle(color = "black")$
  encode(
    x = "x:Q",
    y = "y:Q"
  )

chart_predict <-
  alt$Chart(data_predict)$
  mark_line()$
  encode(
    x = "x:Q",
    y = "y:Q",
    color = "degree:N"
  )

chart <- (chart_data + chart_predict)

chart
```

## Scatter Matrix

[Altair example](https://altair-viz.github.io/gallery/scatter_matrix.html)

#### Data

```{r}
glimpse(vega_data$cars())
```

#### Chart

```{r}
chart <- 
  alt$Chart(vega_data$cars())$
  mark_circle()$
  encode(
    x = alt$X(alt$`repeat`("column"), type = "quantitative"),
    y = alt$Y(alt$`repeat`("row"), type = "quantitative"),
    color = "Origin:N"
  )$
  properties(width = 150, height = 150)$
  `repeat`(
    row = list("Horsepower", "Acceleration", "Miles_per_Gallon"),
    column = list("Miles_per_Gallon", "Acceleration", "Horsepower")
  )$
  interactive()

chart
```



## Scatter Plot with Href


[Altair example](https://altair-viz.github.io/gallery/scatter_href.html)

#### Data

```{r}
glimpse(vega_data$cars())
```

#### Chart

```{r}
chart <- 
  alt$Chart(vega_data$cars())$
  transform_calculate(
    "url", "'https://www.google.com/search?q=' + datum.Name"
  )$
  mark_point()$
  encode(
    x = "Horsepower:Q",
    y = "Miles_per_Gallon:Q",
    color = "Origin:N",
    href = "url:N",
    tooltip = list("Name:N", "url:N")
  )

chart
```



## Scatter Plot with Rolling Mean


[Altair example](https://altair-viz.github.io/gallery/scatter_with_rolling_mean.html)

#### Data

```{r}
glimpse(vega_data$seattle_weather())
```

#### Chart

```{r}
source <- vega_data$seattle_weather()

line <- 
  alt$Chart(source)$
  mark_line(
    color = "red",
    size = 3
  )$
  transform_window(
    rolling_mean = "mean(temp_max)",
    frame = c(-15, 15)
  )$
  encode(
    x = "date:T",
    y = "rolling_mean:Q"
  )


points <- 
  alt$Chart(source)$
  mark_point()$
  encode(
    x = "date:T",
    y = alt$Y(
      "temp_max:Q",
      axis = alt$Axis(title = "Max Temp")
    )
  )

# there seems to be a problem with spec-validation in Altair
# TODO: check using Altair 4.2.0
chart <- points + line
chart$to_json(validate = FALSE) %>% as_vegaspec()
```


## Simple Scatter Plot with Errorbars

[Altair example](https://altair-viz.github.io/gallery/simple_scatter_with_errorbars.html)

#### Data

<details>
<summary>Definition</summary>
```{r}
set.seed(0)
source <- tibble(
  x = 1:5, 
  y = rnorm(5, 10, 0.5),
  yerr = 0.2
)
```
</details>

```{r}
glimpse(source)
```

#### Chart

```{r}
# the base chart
base <- 
  alt$Chart(source)$
  transform_calculate(
    ymin = "datum.y-datum.yerr",
    ymax = "datum.y+datum.yerr"
  )

# generate the points
points <- 
  base$
  mark_point(filled = TRUE, size = 50, color = "black")$
  encode(
    x=alt$X("x", scale = alt$Scale(domain=c(0, 6))),
    y=alt$Y("y", scale = alt$Scale(domain=c(10, 11)))
  )

# generate the error bars
errorbars <-
  base$
  mark_errorbar()$
  encode(
    x = "x",
    y = "ymin:Q",
    y2 = "ymax:Q"
  )

points + errorbars
```


## Simple Scatter Plot with Labels

[Altair example](https://altair-viz.github.io/gallery/scatter_with_labels.html)

Layering can now work with the `+` operator as it does in Python. Additionally, it can be declared as `alt$layer(chart1, chart2)` or as `alt$ChartLayer(layer = list(chart1, chart2))`

#### Data

<details>
<summary>Definition</summary>
```{r}
data <- 
  tibble(
    x = c(1, 3, 5, 7, 9),
    y = c(1, 3, 5, 7, 9),
    label = c("A", "B", "C", "D", "E")
  )
```
</details>

```{r}
glimpse(data)
```

#### Chart

```{r}
bars <- 
  alt$Chart(data)$
  mark_point()$
  encode(
    x = "x:Q",
    y = "y:Q"
  )

text <-
  bars$
  mark_text(align = "left", baseline = "middle", dx = 7)$
  encode(text = "label:N")

chart <- (text + bars)

chart
```

## Stripplot

[Altair example](https://altair-viz.github.io/gallery/stripplot.html)

#### Data

```{r}
glimpse(fromJSON(vega_data$movies$url))
```

#### Chart

```{r}
source <- vega_data$movies$url

stripplot <-
  alt$Chart(source, width = 40)$
  transform_calculate(
    # Generate Gaussian jitter with a Box-Muller transform
    jitter = 'sqrt(-2*log(random()))*cos(2*PI*random())'
  )$
  mark_circle(size = 8)$
  encode(
    x = alt$X(
      'jitter:Q',
      title = " ",
      axis = alt$Axis(
        values = list(0), 
        ticks = TRUE, 
        grid = FALSE, 
        labels = FALSE
      ),
      scale = alt$Scale()
    ),
    y = alt$Y("IMDB_Rating:Q"),
    color = alt$Color("Major_Genre:N", legend = NULL),
    column = alt$Column(
      "Major_Genre:N",
      header = alt$Header(
        labelAngle = -90,
        titleOrient = "top",
        labelOrient = "bottom",
        labelAlign = "right",
        labelPadding = 3
      )
    )
  )$
  configure_facet(spacing = 0)$
  configure_view(stroke = "transparent")

stripplot
```

## Table Bubble Plot (Github Punch Card)

[Altair example](https://altair-viz.github.io/gallery/scatter_with_labels.html)

#### Data

```{r}
glimpse(vega_data$github())
```

#### Chart

```{r}
chart <- 
  alt$Chart(vega_data$github$url)$
  mark_circle()$
  encode(
    x = alt$X("time:O", timeUnit = "hours"),
    y = alt$Y("time:O", timeUnit = "day"),
    size = "sum(count):Q"
  )

chart
```

## Trellis Scatter Plot

[Altair example](https://altair-viz.github.io/gallery/trellis_scatter_plot.html)

This example demonstrates the need for a `ggplot2::facet_wrap()`-like capability, which we understand is coming to Vega-Lite in the no-so-distant future.

#### Data

```{r}
glimpse(vega_data$cars())
```

#### Chart

```{r}
chart <- 
  alt$Chart(vega_data$cars())$
  mark_point()$
  encode(
    x = "Horsepower:Q",
    y = "Miles_per_Gallon:Q",
    row = "Origin:N"
  )

chart
```








