---
title: "Other Charts"
author: "Haley Jeppson"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This document is adapted from the [Other Charts](https://altair-viz.github.io/gallery/index.html#other-charts) section of the Altair Example Gallery.

Our first step is to set up our environment:

```{r}
# devtools::install_github("vegawidget/altair")
library("altair")
library("tibble")
library("dplyr")
library("jsonlite")

vega_data <- import_vega_data()
```

## Bar Chart with Highlighted Segment

[Altair Example](https://altair-viz.github.io/gallery/bar_chart_with_highlighted_segment.html)

#### Data

```{r}
glimpse(vega_data$wheat())
```

#### Chart

```{r}
source <- vega_data$wheat()
threshold <- tibble(threshold = 90)

bars <- 
  alt$Chart(source)$
  mark_bar()$
  encode(
    x = "year:O",
    y = "wheat:Q"
  )

highlight <- 
  alt$Chart(source)$
  mark_bar(color = "#e45755")$
  encode(
    x = "year:O",
    y = "baseline:Q",
    y2 = "wheat:Q"
  )$
  transform_filter("datum.wheat > 90")$
  transform_calculate("baseline", "90")

rule <-
  alt$Chart(threshold)$
  mark_rule()$
  encode(
    y = "threshold:Q"
  )

(bars + highlight + rule)$properties(width = 600)
```

## Becker’s Barley Trellis Plot (wrapped facet)


[Altair Example](https://altair-viz.github.io/gallery/beckers_barley_wrapped_facet.html)


#### Data

```{r}
glimpse(vega_data$barley())
```

#### Chart

```{r}
chart <- 
  alt$Chart(vega_data$barley())$
  mark_point()$
  encode(
    x = alt$X("median(yield)", scale=alt$Scale(zero = FALSE)),
    y = "variety:O",
    color = "year:N",
    facet = "site:O"
  )$
  properties(columns = 2, width = 200, height = 100)

chart
```


## Binned Movie Rating Heatmap

[Altair Example](https://altair-viz.github.io/gallery/binned_heatmap.html)

#### Data

```{r}
glimpse(fromJSON(vega_data$movies$url))
```

#### Chart

```{r}
chart <- 
  alt$Chart(vega_data$movies$url)$
  mark_rect()$
  encode(
    x = alt$X("IMDB_Rating:Q", bin = alt$Bin(maxbins = 60)),
    y = alt$Y("Rotten_Tomatoes_Rating:Q", bin = alt$Bin(maxbins = 40)),
    color = alt$Color(
      "count(IMDB_Rating):Q", 
      scale = alt$Scale(scheme = "greenblue")
    )
  )
chart
```

## Box Plot with Min/Max Whiskers

[Altair example](https://altair-viz.github.io/gallery/boxplot_max_min.html)

> This example shows how to make a basic box plot using US Population data from 2000.

```{r}
glimpse(vega_data$population())
```

#### Chart

```{r}
chart <-
  alt$Chart(vega_data$population$url)$
  mark_boxplot(extent="min-max")$
  encode(
    x = "age:O",
    y = "people:Q"
  )

chart
```

## Candlestick Chart

[Altair example](https://altair-viz.github.io/gallery/candlestick_chart.html)

> A candlestick chart [inspired from Protovis](http://mbostock.github.io/protovis/ex/candlestick.html). This example shows the performance of the Chicago Board Options Exchange Volatility Index (VIX) in the summer of 2009. The thick bar represents the opening and closing prices, while the thin bar shows intraday high and low prices; if the index closed higher on a given day, the bars are colored green rather than red.
#### Data

<details>
<summary>Definition</summary>
```{r}
data <- 
  jsonlite::fromJSON('[
  {
    "date": "01-Jun-2009",
    "open": 28.7,
    "high": 30.05,
    "low": 28.45,
    "close": 30.04,
    "signal": "short",
    "ret": -4.89396411092985
  },
  {
    "date": "02-Jun-2009",
    "open": 30.04,
    "high": 30.13,
    "low": 28.3,
    "close": 29.63,
    "signal": "short",
    "ret": -0.322580645161295
  },
  {
    "date": "03-Jun-2009",
    "open": 29.62,
    "high": 31.79,
    "low": 29.62,
    "close": 31.02,
    "signal": "short",
    "ret": 3.68663594470045
  },
  {
    "date": "04-Jun-2009",
    "open": 31.02,
    "high": 31.02,
    "low": 29.92,
    "close": 30.18,
    "signal": "short",
    "ret": 4.51010886469673
  },
  {
    "date": "05-Jun-2009",
    "open": 29.39,
    "high": 30.81,
    "low": 28.85,
    "close": 29.62,
    "signal": "short",
    "ret": 6.08424336973478
  },
  {
    "date": "08-Jun-2009",
    "open": 30.84,
    "high": 31.82,
    "low": 26.41,
    "close": 29.77,
    "signal": "short",
    "ret": 1.2539184952978
  },
  {
    "date": "09-Jun-2009",
    "open": 29.77,
    "high": 29.77,
    "low": 27.79,
    "close": 28.27,
    "signal": "short",
    "ret": -5.02431118314424
  },
  {
    "date": "10-Jun-2009",
    "open": 26.9,
    "high": 29.74,
    "low": 26.9,
    "close": 28.46,
    "signal": "short",
    "ret": -5.46623794212217
  },
  {
    "date": "11-Jun-2009",
    "open": 27.36,
    "high": 28.11,
    "low": 26.81,
    "close": 28.11,
    "signal": "short",
    "ret": -8.3743842364532
  },
  {
    "date": "12-Jun-2009",
    "open": 28.08,
    "high": 28.5,
    "low": 27.73,
    "close": 28.15,
    "signal": "short",
    "ret": -5.52763819095477
  },
  {
    "date": "15-Jun-2009",
    "open": 29.7,
    "high": 31.09,
    "low": 29.64,
    "close": 30.81,
    "signal": "long",
    "ret": 3.4920634920635
  },
  {
    "date": "16-Jun-2009",
    "open": 30.81,
    "high": 32.75,
    "low": 30.07,
    "close": 32.68,
    "signal": "short",
    "ret": 0.155038759689914
  },
  {
    "date": "17-Jun-2009",
    "open": 31.19,
    "high": 32.77,
    "low": 30.64,
    "close": 31.54,
    "signal": "short",
    "ret": 5.82822085889571
  },
  {
    "date": "18-Jun-2009",
    "open": 31.54,
    "high": 31.54,
    "low": 29.6,
    "close": 30.03,
    "signal": "short",
    "ret": 8.17610062893082
  },
  {
    "date": "19-Jun-2009",
    "open": 29.16,
    "high": 29.32,
    "low": 27.56,
    "close": 27.99,
    "signal": "short",
    "ret": 8.59872611464968
  },
  {
    "date": "22-Jun-2009",
    "open": 30.4,
    "high": 32.05,
    "low": 30.3,
    "close": 31.17,
    "signal": "short",
    "ret": 15.4907975460123
  },
  {
    "date": "23-Jun-2009",
    "open": 31.3,
    "high": 31.54,
    "low": 27.83,
    "close": 30.58,
    "signal": "short",
    "ret": 11.7370892018779
  },
  {
    "date": "24-Jun-2009",
    "open": 30.58,
    "high": 30.58,
    "low": 28.79,
    "close": 29.05,
    "signal": "long",
    "ret": -10.4234527687296
  },
  {
    "date": "25-Jun-2009",
    "open": 29.45,
    "high": 29.56,
    "low": 26.3,
    "close": 26.36,
    "signal": "long",
    "ret": 0
  },
  {
    "date": "26-Jun-2009",
    "open": 27.09,
    "high": 27.22,
    "low": 25.76,
    "close": 25.93,
    "signal": "long",
    "ret": 0
  },
  {
    "date": "29-Jun-2009",
    "open": 25.93,
    "high": 27.18,
    "low": 25.29,
    "close": 25.35,
    "signal": "long",
    "ret": 5.26315789473684
  },
  {
    "date": "30-Jun-2009",
    "open": 25.36,
    "high": 27.38,
    "low": 25.02,
    "close": 26.35,
    "signal": "long",
    "ret": 6.73758865248228
  }
]')
```
</details>

```{r}
glimpse(data)
```

#### Chart

```{r}
open_close_color <- 
  alt$condition(
    "datum.open < datum.close",
    alt$value("#06982d"),
    alt$value("#ae1325")
  )

rule <- 
  alt$Chart(data)$
  mark_rule()$
  encode(
    alt$X(
      "date:T",
      timeUnit = "yearmonthdate",
      scale = alt$Scale(
        domain = list(
          list(month= 5, date= 31, year= 2009),
          list(month= 7, date= 1, year= 2009)
        )
      ),
      axis = alt$Axis(format = "%m/%d", title = "Date in 2009")
    ),
    alt$Y(
      "low",
      scale = alt$Scale(zero = FALSE),
      axis = alt$Axis(title = "Price")
    ),
    alt$Y2("high"),
    color = open_close_color
  )

bar <- 
  alt$Chart(data)$
  mark_bar()$
  encode(
    alt$X("date:T", timeUnit = "yearmonthdate"),
    y = "open",
    y2 = "close",
    color = open_close_color
  )

chart <- (rule + bar)
chart
```

## Error Bar with Standard Deviation


[Altair Example](https://altair-viz.github.io/gallery/errorbars_with_std.html)

#### Data

```{r}
glimpse(vega_data$barley())
```

#### Chart

```{r}
source <- vega_data$barley()

error_bars <-
  alt$Chart(source)$
  mark_errorbar(extent = "stdev")$
  encode( 
    x = alt$X("yield:Q", scale = alt$Scale(zero = FALSE)),
    y = alt$Y("variety:N")
  )

points <-
  alt$Chart(source)$
  mark_point(filled = TRUE, color = "black")$
  encode(
    x = alt$X("yield:Q", aggregate = "mean"),
    y = alt$Y("variety:N")
  )

error_bars + points
```

## Error Bars showing Confidence Interval


[Altair Example](https://altair-viz.github.io/gallery/errorbars_with_ci.html)

> This example shows how to show error bars using confidence intervals. The confidence intervals are computed internally in vega by a non-parametric [bootstrap of the mean](https://github.com/vega/vega-statistics/blob/master/src/bootstrapCI.js).

#### Data

```{r}
glimpse(vega_data$barley())
```

#### Chart

```{r}
source <- vega_data$barley()

error_bars <-
  alt$Chart(source)$
  mark_errorbar(extent = "ci")$
  encode(
    x = alt$X("yield:Q", scale = alt$Scale(zero = FALSE)),
    y = alt$Y("variety:N")
  )

points <-
  alt$Chart(source)$
  mark_point(filled = TRUE, color = "black")$
  encode(
    x = alt$X("yield:Q", aggregate = "mean"),
    y = alt$Y("variety:N")
  )

error_bars + points
```

## Facetted Scatterplot with marginal histograms

[Altair Example](https://altair-viz.github.io/gallery/scatter_marginal_hist.html)

There is a slight difference from the Altair example in the `yscale`, to get everything to line-up correctly.

#### Data

```{r}
glimpse(vega_data$iris())
```

#### Chart

```{r}
iris <- vega_data$iris()

xscale <- alt$Scale(domain = list(4.0, 8.0))
yscale <- alt$Scale(domain = list(1.8, 4.6))

blank_axis <- alt$Axis(title = "")

points <- 
  alt$Chart(iris)$
  mark_circle()$
  encode(
    alt$X("sepalLength", scale = xscale),
    alt$Y("sepalWidth", scale = yscale),
    color = "species"
  )

top_hist <-
  alt$Chart(iris)$
  mark_area(opacity = 0.3, interpolate = "step")$
  encode(
    x = alt$X(
      "sepalLength:Q",
      # when using bins, the axis scale is set through
      # the bin extent, so we do not specify the scale here
      # (which would be ignored anyway)
      bin = alt$Bin(maxbins = 20, extent = xscale$domain),
      stack = NULL,
      axis = blank_axis
    ),
    y = alt$Y("count(sepalLength)", stack = NULL, axis = blank_axis),
    color = alt$Color("species:N")
  )$
  properties(height = 60)

right_hist <- 
  alt$Chart(iris)$
  mark_area(opacity = 0.3, interpolate = "step")$
  encode(
    x = alt$X("count(sepalWidth)", stack = NULL, axis = blank_axis),
    y = alt$Y(
      "sepalWidth:Q",
      bin = alt$Bin(maxbins = 20, extent = yscale$domain),
      stack = NULL,
      axis = blank_axis
    ),
    color = alt$Color("species:N")
  )$
  properties(width = 60)

chart <- (top_hist & (points | right_hist))

chart
```

## Gantt Chart

[Altair example](https://altair-viz.github.io/gallery/gantt_chart.html)

#### Data

<details>
<summary>Definition</summary>
```{r}
data <- 
  fromJSON('[
    {"task": "A", "start": 1, "end": 3},
    {"task": "B", "start": 3, "end": 8},
    {"task": "C", "start": 8, "end": 10}
  ]')
```
</details>

```{r}
glimpse(data)
```

#### Chart

```{r}
chart <- 
  alt$Chart(data)$
  mark_bar()$
  encode(
    x = "start",
    x2 = "end",
    y = "task"
  )

chart
```


## Isotype Grid


[Altair Example](https://altair-viz.github.io/gallery/isotype_grid.html)

#### Data

<details>
<summary>Definition</summary>
```{r}
data <- tibble(id = 1:100)
person = c("M1.7 -1.7h-0.8c0.3 -0.2 0.6 -0.5 0.6 -0.9c0 -0.6 -0.4 -1 -1 -1c-0.6 0 -1 0.4 -1 1c0 0.4 0.2 0.7 0.6 0.9h-0.8c-0.4 0 -0.7 0.3 -0.7 0.6v1.9c0 0.3 0.3 0.6 0.6 0.6h0.2c0 0 0 0.1 0 0.1v1.9c0 0.3 0.2 0.6 0.3 0.6h1.3c0.2 0 0.3 -0.3 0.3 -0.6v-1.8c0 0 0 -0.1 0 -0.1h0.2c0.3 0 0.6 -0.3 0.6 -0.6v-2c0.2 -0.3 -0.1 -0.6 -0.4 -0.6z")
```
</details>

```{r}
glimpse(data)
```

#### Chart

```{r}
chart <-
  alt$Chart(data)$
  transform_calculate(row = "ceil(datum.id/10)")$
  transform_calculate(col = "datum.id - datum.row*10")$
  mark_point(filled = TRUE, size = 50)$
  encode(
    x = alt$X("col:O", axis = NULL),
    y = alt$Y("row:O", axis = NULL),
    shape = alt$ShapeValue(person)
  )$
  properties(width = 400, height = 400)$
  configure_view(strokeWidth = 0)

chart
```


## Multiple Marks

[Altair Example](https://altair-viz.github.io/gallery/multiple_marks.html)

#### Data

```{r}
glimpse(vega_data$stocks())
```

#### Chart

```{r}
stocks <- vega_data$stocks()

chart <- 
  alt$Chart(stocks)$
  mark_line(point = TRUE)$
  encode(
    x = "date:T",
    y = "price:Q",
    color = "symbol:N"
  )

chart
```

## Normalized Parallel Coordinates Example


[Altair Example](https://altair-viz.github.io/gallery/normed_parallel_coordinates.html)

> A [Parallel Coordinates](https://en.wikipedia.org/wiki/Parallel_coordinates) chart is a chart that lets you visualize the individual data points by drawing a single line for each of them.
>
> Such a chart can be created in Altair by first transforming the data into a suitable representation.
>
> This example shows a modified parallel coordinates chart with the Iris dataset, where the y-axis shows the value after min-max rather than the raw value. It’s a simplified Altair version of [the VegaLite version](https://vega.github.io/vega-lite/examples/parallel_coordinate.html).

#### Data

```{r}
glimpse(vega_data$iris())
```

#### Chart

```{r}
source <- vega_data$iris()

chart <-
  alt$Chart(source)$
  transform_window(index="count()")$
  transform_fold(
    list("petalLength", "petalWidth", "sepalLength", "sepalWidth")
  )$
  transform_joinaggregate(
    groupby = list("key"),
    min = "min(value)",
    max = "max(value)"
  )$
  transform_calculate(
    minmax_value = "(datum.value - datum.min) / (datum.max - datum.min)",
    mid = "(datum.min + datum.max) / 2"
  )$
  mark_line()$
  encode(
    x = "key:N",
    y = "minmax_value:Q",
    color = "species:N",
    detail = "index:N",
    opacity = alt$value(0.5)
  )$properties(width = 500)

chart
```

## Parallel Coordinates Example


[Altair Example](https://altair-viz.github.io/gallery/parallel_coordinates.html)

> A [Parallel Coordinates](https://en.wikipedia.org/wiki/Parallel_coordinates) chart is a chart that lets you visualize the individual data points by drawing a single line for each of them. Such a chart can be created in Altair by first transforming the data into a suitable representation. This example shows a parallel coordinates chart with the Iris dataset.

#### Data

```{r}
glimpse(vega_data$iris())
```

#### Chart

```{r}
source <- vega_data$iris()

chart <-
  alt$Chart(source)$
  transform_window(index = "count()")$
  transform_fold(
    list("petalLength", "petalWidth", "sepalLength", "sepalWidth")
  )$
  mark_line()$
  encode(
    x = "key:N",
    y = "value:Q",
    color = "species:N",
    detail = "index:N",
    opacity = alt$value(0.5)
  )$
  properties(width = 500)

chart
```


## Ranged Dot Plot

[Altair Example](https://altair-viz.github.io/gallery/ranged_dot_plot.html)

> This example shows a ranged dot plot that uses *layer* to convey changing life expectancy for the five most populous countries (between 1955 and 2000).

#### Data

Cannot get the Vega-Lite data-layer to work.

<details>
<summary>Definition</summary>
```{r}
data <- fromJSON(vega_data$countries$url)
data <- 
  data %>%
  select(-`_comment`) %>%
  filter(
    country %in% c("China", "India", "United States", "Indonesia", "Brazil"),
    year %in% c(1955, 2000)
  ) 
```
</details>

```{r}
glimpse(data)
```

#### Chart

```{r}
# Line between life expectancy in 1955 & 2000
chart_line <-
  alt$Chart(data)$
  mark_line(color = "#db646f")$
  encode(
    x = "life_expect:Q",
    y = "country:N",
    detail = "country:N"
  )

# Points for life expectancy in 1955 & 2000
chart_point <- 
  alt$Chart(data)$
  mark_point(size = 100, opacity = 1, filled = TRUE)$
  encode(
    x = "life_expect:Q",
    y = "country:N",
    color = alt$Color(
      "year:O",
      scale = alt$Scale(
        domain = list("1955", "2000"), 
        range = list("#e6959c", "#911a24")
      )
    )
  )$interactive()

# Compose charts, add data and transformations
chart <- 
  (chart_line + chart_point)$
  transform_filter(
    filter = list(
      field = "country",
      oneOf = list("China", "India", "United States", "Indonesia", "Brazil")
    )
  )$
  transform_filter(
    filter = list(field = "year", oneOf = list(1955, 2000))
  )

chart
```

## Ridgeline plot Example


[Altair Example](https://altair-viz.github.io/gallery/ridgeline_plot.html)

> A [Ridgeline plot](https://serialmentor.com/blog/2017/9/15/goodbye-joyplots) chart is a chart that lets you visualize distribution of a numeric value for several groups.
>
> Such a chart can be created in Altair by first transforming the data into a suitable representation.

#### Data

```{r}
glimpse(vega_data$seattle_weather())
```

#### Chart

```{r}
source <- vega_data$seattle_weather()

step <- 20
overlap <- 1

chart <-alt$Chart(source)$
  transform_timeunit(Month = "month(date)")$
  transform_joinaggregate(
    mean_temp = "mean(temp_max)", 
    groupby = list("Month")
  )$
  transform_bin(list("bin_max", "bin_min"), "temp_max")$
  transform_aggregate(
    value = "count()", 
    groupby = list("Month", "mean_temp", "bin_min", "bin_max")
  )$
  transform_impute(
    impute = "value", 
    groupby = list("Month", "mean_temp"), 
    key = "bin_min", 
    value = 0
  )$
  mark_area(
    interpolate = "monotone",
    fillOpacity = 0.8,
    stroke = "lightgray",
    strokeWidth = 0.5
  )$encode(
    alt$X("bin_min:Q", bin = "binned", title = "Maximum Daily Temperature (C)"),
    alt$Y(
      "value:Q",
      scale = alt$Scale(range = list(step, -step * overlap)),
      axis = NULL
    ),
    alt$Fill(
      "mean_temp:Q",
      legend = NULL,
      scale = alt$Scale(domain = list(30, 5), scheme = "redyellowblue")
    ),
    alt$Row(
      "Month:T",
      title = NULL,
      header = alt$Header(labelAngle = 0, labelAlign = "right", format = "%B")
    )
  )$
  properties(bounds ="flush", title = "Seattle Weather", height = step)$
  configure_facet(spacing = 0)$
  configure_view(stroke = NULL)$
  configure_title(anchor = "end")

chart
```


## Sorted Error Bars showing Confidence Interval


[Altair Example](https://altair-viz.github.io/gallery/sorted_error_bars_with_ci.html)

> This example shows how to show error bars using confidence intervals, while also sorting the y-axis based on x-axis values.

#### Data

```{r}
glimpse(vega_data$barley())
```

#### Chart

```{r}
source <- vega_data$barley()

points <-
  alt$Chart(source)$
  mark_point(filled = TRUE, color = "black")$
  encode(
    x=alt$X("mean(yield):Q", title="Barley Yield"),
    y=alt$Y( 
      "variety",
      sort = alt$EncodingSortField(
        field = "yield",
        op = "mean",
        order = "descending"
      )
    )
  )$properties(width = 400, height = 250)

error_bars <-
  points$
  mark_rule()$
  encode(
    x = "ci0(yield)",
    x2 = "ci1(yield)"
  )

points + error_bars
```



## Steam and Leaf Plot

[Altair Example](https://altair-viz.github.io/gallery/stem_and_leaf.html)

#### Data

<details>
<summary>Definition</summary>
```{r}
data <- 
  tibble(
    sample = rnorm(100, 50, 15) %>% round() %>% as.integer(),
    stem = (sample / 10L) %>% as.integer(),
    leaf = sample %% 10L
  ) %>%
  arrange(sample) %>%
  group_by(stem) %>%
  mutate(position = seq_along(leaf)) %>%
  ungroup()
```
</details>

```{r}
glimpse(data)
```

#### Chart

```{r}
chart <- 
  alt$Chart(data)$
  mark_text(align = "left", baseline = "middle", dx = -5)$
  encode(
    x = alt$X(
      "position:Q",
      axis = alt$Axis(title="", ticks = FALSE, labels = FALSE, grid = FALSE)
    ),
    y = alt$Y("stem:N", axis = alt$Axis(title = "", tickSize = 0)),
    text = "leaf:N"
  )$
  configure_axis(labelFontSize = 20)$
  configure_text(fontSize = 20)

chart
```

## Text over a Heatmap


[Altair Example](https://altair-viz.github.io/gallery/layered_heatmap_text.html)

> An example of a layered chart of text over a heatmap using the cars dataset.

#### Data

```{r}
glimpse(vega_data$cars())
```

#### Chart

```{r}
source <- vega_data$cars()

# Configure common options
base <-
  alt$Chart(source)$
  transform_aggregate(
    num_cars="count()",
    groupby=list("Origin", "Cylinders")
  )$encode(
    alt$X("Cylinders:O", scale=alt$Scale(paddingInner=0)),
    alt$Y("Origin:O", scale=alt$Scale(paddingInner=0))
  )

# Configure heatmap
heatmap <- base$
  mark_rect()$
  encode(
    color=alt$Color("num_cars:Q",
                    scale=alt$Scale(scheme="viridis"),
                    legend=alt$Legend(direction="horizontal")
    )
  )

# Configure text
text <- base$
  mark_text(baseline="middle")$
  encode(
    text="num_cars:Q",
    color=alt$condition(
      "datum.num_cars > 100",
      alt$value("black"),
      alt$value("white")
    )
  )

# Draw the chart
heatmap + text
```



## Violinplot

[Altair Example](https://altair-viz.github.io/gallery/violine_plot.html)

> This example shows how to make a kind of a Violinplot.


#### Data

```{r}
glimpse(vega_data$cars())
```

#### Chart

```{r}
source <- vega_data$cars()

chart <-
  alt$Chart(source)$
  transform_filter(
    "datum.Miles_per_Gallon > 0"
  )$transform_bin(
    list("bin_max", "bin_min"), field="Miles_per_Gallon", bin=alt$Bin(maxbins=20)
  )$transform_calculate(
    binned="(datum.bin_max + datum.bin_min) / 2"
  )$transform_aggregate(
    value_count="count()", groupby=list("Origin", "binned")
  )$transform_impute(
    impute="value_count", groupby=list("Origin"), key="binned", value=0
  )$mark_area(
    interpolate="monotone",
    orient="horizontal"
  )$encode(
    x=alt$X(
      "value_count:Q",
      title=NULL,
      stack="center",
      axis=alt$Axis(labels=FALSE, values=list(0),grid=FALSE, ticks=TRUE)
    ),
    y=alt$Y("binned:Q", bin="binned", title="Miles per Gallon"),
    color=alt$Color("Origin:N", legend=NULL),
    column=alt$Column(
      "Origin:N",
      header=alt$Header(
        titleOrient="bottom",
        labelOrient="bottom",
        labelPadding=0
      )
    )
  )$properties(
    width=80
  )$configure_facet(
    spacing=0
  )$configure_view(
    stroke=NULL
  )

chart
```
