---
title: "Simple Charts"
author: "Haley Jeppson"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This document is adapted from the [Simple Charts](https://altair-viz.github.io/gallery/#simple-charts) section of the Altair Example Gallery.

Our first step is to set up our environment:

```{r message=FALSE}
library("altair")
library("reticulate")
library("tibble")
library("tidyr")
library("dplyr")

vega_data <- import_vega_data()
```

## Simple Bar Chart

[Altair example](https://altair-viz.github.io/gallery/simple_bar_chart.html)  

#### Data

<details>
<summary>Definition</summary>
```{r}
data <-  
  tibble(
    a= c("A", "B", "C", "D", "E", "F", "G", "H", "I"),
    b =c(28, 55, 43, 91, 81, 53, 19, 87, 52)
  )
```
</details>

```{r}
glimpse(data)
```

#### Chart

```{r}
chart <- 
  alt$Chart(data)$
  mark_bar()$
  encode(
    x = "a",
    y = "b"
  )

chart
```

## Simple Heatmap

[Altair example](https://altair-viz.github.io/gallery/simple_heatmap.html)  

#### Data

<details>
<summary>Definition</summary>
```{r}
# Compute x^2 + y^2 across a 2D grid
data <- 
  crossing(x = seq(-5, 5), y = seq(-5, 5)) %>%
  mutate(z = x^2 + y^2)
```
</details>

```{r}
glimpse(data)
```

#### Chart

```{r}
chart <- 
  alt$Chart(data)$
  mark_rect()$
  encode(
    x = "x:O",
    y = "y:O",
    color = "z:Q"
  )

chart
```

## Simple Histogram

[Altair example](https://altair-viz.github.io/gallery/simple_histogram.html)  

> This example shows how to make a basic histogram, based on the [vega-lite docs](https://vega.github.io/vega-lite/examples/histogram.html). 

#### Data

```{r}
movies <- jsonlite::fromJSON("https://vega.github.io/vega-datasets/data/movies.json")
glimpse(movies)
```

#### Chart

```{r}
chart <- 
  alt$Chart(movies)$
  mark_bar()$
  encode(
    alt$X("IMDB Rating:Q", bin = TRUE), 
    y = "count()"
  )

chart
```

## Simple Line Chart

[Altair example](https://altair-viz.github.io/gallery/simple_line_chart.html)

#### Data

<details>
<summary>Definition</summary>
```{r}
data <- 
  data_frame(
    x = 0:100,
    sinx = sin(x / 5)
  ) 
```
</details>

```{r}
glimpse(data)
```

#### Chart

```{r}
chart <- 
  alt$Chart(data)$
  mark_line()$
  encode(
    x = "x",
    y = "sinx"
  )

chart
```

## Simple Scatter Plot

[Altair example](https://altair-viz.github.io/gallery/scatter_tooltips.html)

#### Data

```{r}
glimpse(vega_data$cars())
```

#### Chart

```{r}
chart <- 
  alt$Chart(vega_data$cars())$
  mark_circle()$
  encode(
    x = "Horsepower",
    y = "Miles_per_Gallon",
    color = "Origin",
    tooltip = c("Name", "Horsepower", "Miles_per_Gallon", "Origin")
  )$
  interactive()

chart
```

## Simple Stacked Area Chart

[Altair example](https://altair-viz.github.io/gallery/simple_stacked_area_chart.html)

#### Data

```{r}
glimpse(vega_data$iowa_electricity())
```

#### Chart

```{r}
chart <- 
  alt$Chart(vega_data$iowa_electricity())$
  mark_area()$
  encode(
    x = "year:T",
    y = "net_generation:Q",
    color = "source:N"
  )

chart
```

## Strip Plot

[Altair example](https://altair-viz.github.io/gallery/strip_plot.html)

#### Data

```{r}
glimpse(vega_data$cars())
```

#### Chart

```{r}
chart <- 
  alt$Chart(vega_data$cars())$
  mark_tick()$
  encode(
    x = "Horsepower:Q",
    y = "Cylinders:O"
  )

chart
```







