---
title: "Line Charts"
author: "Haley Jeppson"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This document is adapted from the [Bar Charts](https://altair-viz.github.io/gallery/index.html#bar-charts) section of the Altair Example Gallery.

Our first step is to set up our environment:

```{r}
# devtools::install_github("vegawidget/altair")
library("altair")
library("tibble")
library("jsonlite")
library("dplyr")
library("readr")

vega_data <- import_vega_data()
```

## Cumulative Wikipedia Donations

[Altair example](https://altair-viz.github.io/gallery/cumulative_wiki_donations.html)

> This chart shows cumulative donations to Wikipedia over the past 10 years. Inspired by this [Reddit post](https://www.reddit.com/r/dataisbeautiful/comments/7guwd0/cumulative_wikimedia_donations_over_the_past_10/) but using lines instead of areas.

#### Data

```{r}
data <- read_csv(curl::curl("https://frdata.wikimedia.org/donationdata-vs-day.csv"))

glimpse(data)
```

#### Chart

```{r}
chart <- 
  alt$Chart(r_to_py(data))$
  mark_line()$
  encode(
    x = alt$X(
      "date:T", 
      timeUnit = "monthdate",
      axis = alt$Axis(format = "%B", title = "Month")
    ),
    y = alt$Y(
      "max(ytdsum):Q", 
      stack = NULL,
      axis = alt$Axis(title = "Cumulative Donations")
    ),
    color = alt$Color(
      "date:O", 
      timeUnit = "year", 
      legend = alt$Legend(title = "Year")
    ),
    order = alt$Order("data:O", timeUnit= "year")
  )

chart
```

## Falkensee

[Altair example](https://altair-viz.github.io/gallery/falkensee.html)

#### Data

<details>
<summary>Definition</summary>
```{r}
source <- fromJSON('[
    {"year": "1875", "population": 1309},
    {"year": "1890", "population": 1558},
    {"year": "1910", "population": 4512},
    {"year": "1925", "population": 8180},
    {"year": "1933", "population": 15915},
    {"year": "1939", "population": 24824},
    {"year": "1946", "population": 28275},
    {"year": "1950", "population": 29189},
    {"year": "1964", "population": 29881},
    {"year": "1971", "population": 26007},
    {"year": "1981", "population": 24029},
    {"year": "1985", "population": 23340},
    {"year": "1989", "population": 22307},
    {"year": "1990", "population": 22087},
    {"year": "1991", "population": 22139},
    {"year": "1992", "population": 22105},
    {"year": "1993", "population": 22242},
    {"year": "1994", "population": 22801},
    {"year": "1995", "population": 24273},
    {"year": "1996", "population": 25640},
    {"year": "1997", "population": 27393},
    {"year": "1998", "population": 29505},
    {"year": "1999", "population": 32124},
    {"year": "2000", "population": 33791},
    {"year": "2001", "population": 35297},
    {"year": "2002", "population": 36179},
    {"year": "2003", "population": 36829},
    {"year": "2004", "population": 37493},
    {"year": "2005", "population": 38376},
    {"year": "2006", "population": 39008},
    {"year": "2007", "population": 39366},
    {"year": "2008", "population": 39821},
    {"year": "2009", "population": 40179},
    {"year": "2010", "population": 40511},
    {"year": "2011", "population": 40465},
    {"year": "2012", "population": 40905},
    {"year": "2013", "population": 41258},
    {"year": "2014", "population": 41777}
  ]')

source2 <- fromJSON('[
    {
      "start": "1933",
      "end": "1945",
      "event": "Nazi Rule"
    },
    {
      "start": "1948",
      "end": "1989",
      "event": "GDR (East Germany)"
    }
  ]')
```
</details>

```{r}
glimpse(source)
```

```{r}
glimpse(source2)
```

#### Chart

```{r}
source <- r_to_py(source)
source2 <- r_to_py(source2)

line <-
  alt$Chart(source)$
  mark_line(color="#333")$
  encode(
    x = alt$X("year:T", axis = alt$Axis(format = "%Y")),
    y = "population"
  )$
  properties(
    width = 600,
    height = 400
  )

point <- line$mark_point(color="#333")

rect <- 
  alt$Chart(source2)$
  mark_rect()$
  encode(
    x = "start:T",
    x2 = "end:T",
    color = "event:N"
  )

chart <- (rect + line + point)

chart
```

## Line Chart with Layered Aggregates


[Altair example](https://altair-viz.github.io/gallery/layer_line_color_rule.html)

> This example shows how to make a multi series line chart of the daily closing stock prices for AAPL, AMZN, GOOG, IBM, and MSFT between 2000 and 2010, along with a layered rule showing the average values.

#### Data

```{r}
glimpse(vega_data$stocks())
```

#### Chart

```{r}
stocks <- r_to_py(vega_data$stocks())

line <- 
  alt$Chart(stocks)$
  mark_line()$
  encode(
    x = "date",
    y = "price",
    color = "symbol"
  )$
  properties(
    width = 600,
    title = "Daily closing prices with their aggregate prices"
  )$
  interactive(bind_y=FALSE)

rule <- 
  alt$Chart(stocks)$
  mark_rule()$
  encode(
    y = alt$Y("average(price)"),
    color = "symbol",
    size = alt$value(2)
  )

chart <- (line + rule)

chart
```

## Line Chart with Percent axis


[Altair example](https://altair-viz.github.io/gallery/line_percent.html)

#### Data

```{r}
jobs <- vega_data$jobs() %>% sample_n(5000)

glimpse(jobs)
```

#### Chart

```{r}
source = r_to_py(jobs)

chart <- 
  alt$Chart(source)$
  mark_line()$
  encode(
    x = alt$X("year:O"),
    y = alt$Y("perc:Q", axis = alt$Axis(format = "%")),
    color = "sex:N"
  )$
  properties(title = "Percent of work-force working as Welders")$
  transform_filter("datum.job == 'Welder'")

chart
```

## Line chart with Confidence Interval Band

[Altair example](https://altair-viz.github.io/gallery/line_with_ci.html)

> This example shows how to make a line chart with a bootstrapped 95% confidence interval band.

#### Data

```{r}
glimpse(vega_data$cars())
```

#### Chart

```{r}
cars <- r_to_py(vega_data$cars())

line <- 
  alt$Chart(cars)$
  mark_line()$
  encode(
    x = "Year",
    y = "mean(Miles_per_Gallon)"
  )

confidence_interval <- 
  alt$Chart(cars)$
  mark_area(opacity=0.3)$
  encode(
    x = "Year",
    y = alt$Y(
      "ci0(Miles_per_Gallon)", 
      axis = alt$Axis(title = "Miles/Gallon")
    ),
    y2 = "ci1(Miles_per_Gallon)"
  )

chart <- (confidence_interval + line)

chart
```

## Multi Series Line Chart


[Altair example](https://altair-viz.github.io/gallery/multi_series_line.html)


#### Data

```{r}
glimpse(vega_data$stocks())
```

#### Chart

Had to make sure to delare that the `date` variable is of type `"temporal"`

```{r}
chart <- 
  alt$Chart(r_to_py(vega_data$stocks()))$
  mark_line()$
  encode(
    x = "date:T",
    y = "price",
    color = "symbol"
  )

chart
```

## Slope Graph


[Altair example](https://altair-viz.github.io/gallery/slope_graph.html)


#### Data

The year here is stored by pandas as an integer. When treating columns as dates, it is best to use either a string representation or a datetime representation.

```{r}
barley <- vega_data$barley()
barley$year <- as.character(barley$year)

glimpse(barley)
```

#### Chart

```{r}
chart <- 
  alt$Chart(r_to_py(barley))$
  mark_line()$
  encode(
    x = "year",
    y = "median(yield)",
    color = "site"
  )

chart
```

## Step Chart


[Altair example](https://altair-viz.github.io/gallery/step_chart.html)

> This example shows Google`s stock price over time. This uses the `step-after` interpolation scheme. The full list of interpolation options includes `linear`, `linear-closed`, `step`, `step-before`, `step-after`, `basis`, `basis-open`, `basis-closed`, `cardinal`, `cardinal-open`, `cardinal-closed`, `bundle`, and `monotone`.

#### Data

```{r}
glimpse(vega_data$stocks())
```

#### Chart

Need to specify the `transform_filter` as a string?

```{r}
chart <- 
  alt$Chart(r_to_py(vega_data$stocks()))$
  mark_line(interpolate = "step-after")$
  encode(
    x = "date",
    y = "price"
  )$
  transform_filter("datum.symbol == 'GOOG'")

chart
```

