---
title: "Other Charts"
author: "Haley Jeppson"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This document is adapted from the [Other Charts](https://altair-viz.github.io/gallery/index.html#other-charts) section of the Altair Example Gallery.

Our first step is to set up our environment:

```{r}
# devtools::install_github("vegawidget/altair")
library("altair")
library("tibble")
library("dplyr")

vega_data <- import_vega_data()
```

## Binned Movie Rating Heatmap

[Altair Example](https://altair-viz.github.io/gallery/binned_heatmap.html)

#### Data

```{r}
glimpse(vega_data$movies())
```

#### Chart

```{r}
chart <- 
  alt$Chart(r_to_py(vega_data$movies()))$
  mark_rect()$
  encode(
    x = alt$X("IMDB_Rating:Q", bin = alt$Bin(maxbins = 60)),
    y = alt$Y("Rotten_Tomatoes_Rating:Q", bin=alt$Bin(maxbins = 40)),
    color = alt$Color(
      "count(IMDB_Rating):Q", 
      scale = alt$Scale(scheme = "greenblue")
    )
  )

chart
```

## Facetted Scatterplot with marginal histograms

[Altair Example](https://altair-viz.github.io/gallery/scatter_marginal_hist.html)

There is a slight difference from the Altair example in the `yscale`, to get everything to line-up correctly.

TODO - note this at ALtair.

#### Data

```{r}
glimpse(vega_data$iris())
```

#### Chart

```{r}
iris <- r_to_py(vega_data$iris())

xscale <- alt$Scale(domain = list(4.0, 8.0))
yscale <- alt$Scale(domain = list(1.8, 4.6))

blank_axis <- alt$Axis(title = "")

points <- 
  alt$Chart(iris)$
  mark_circle()$
  encode(
    alt$X("sepalLength", scale = xscale),
    alt$Y("sepalWidth", scale = yscale),
    color = "species"
  )

top_hist <-
  alt$Chart(iris)$
  mark_area(opacity = 0.3, interpolate = "step")$
  encode(
    x = alt$X("sepalLength:Q",
          # when using bins, the axis scale is set through
          # the bin extent, so we do not specify the scale here
          # (which would be ignored anyway)
          bin = alt$Bin(maxbins = 20, extent = xscale$domain),
          stack = NULL,
          axis = blank_axis
        ),
    y = alt$Y("count(sepalLength)", stack = NULL, axis = blank_axis),
    color = alt$Color("species:N")
  )$
  properties(height=60)

right_hist <- 
  alt$Chart(iris)$
  mark_area(opacity= 0.3, interpolate= "step")$
  encode(
    x = alt$X("count(sepalWidth)", stack=NULL, axis=blank_axis),
    y = alt$Y(
      "sepalWidth:Q",
      bin = alt$Bin(maxbins = 20, extent = yscale$domain),
      stack = NULL,
      axis = blank_axis
    ),
    color = alt$Color("species:N")
  )$
  properties(width=60)

chart <- (top_hist & (points | right_hist))

chart
```

## Layering text over a heatmap

[Altair Example](https://altair-viz.github.io/gallery/layered_heatmap_text.html)

#### Data

```{r}
glimpse(vega_data$cars())
```

#### Chart

```{r}
cars <- r_to_py(vega_data$cars())

heatmap <- 
  alt$Chart(cars)$
  mark_rect()$
  encode(
    x = alt$X("Cylinders:O", scale = alt$Scale(paddingInner = 0)),
    y = alt$Y("Origin:O", scale = alt$Scale(paddingInner = 0)),
    color= "count(Cylinders)"
  )

text <- 
  alt$Chart(cars)$
  mark_text(baseline="middle")$
  encode(
    x = "Cylinders:O",
    y = "Origin:O",
    text = "count(Cylinders)",
    color = alt$condition(
      "datum['count_*'] > 100",
      alt$value("black"),
      alt$value("white")
    )
  )

chart <- (heatmap + text)

chart
```

## Multiple Marks

[Altair Example](https://altair-viz.github.io/gallery/multiple_marks.html)

#### Data

```{r}
glimpse(vega_data$stocks())
```

#### Chart

```{r}
stocks <- r_to_py(vega_data$stocks())

chart <- 
  alt$LayerChart(stocks)$
  encode(
    x = "date:T",
    y = "price:Q",
    color = "symbol:N"
  )$add_layers(
    alt$Chart()$mark_point(),
    alt$Chart()$mark_line()
  )

chart
```

## Ranged Dot Plot

[Altair Example](https://altair-viz.github.io/gallery/ranged_dot_plot.html)

> This example shows a ranged dot plot that uses *layer* to convey changing life expectancy for the five most populous countries (between 1955 and 2000).

#### Data

Cannot get the Vega-Lite data-layer to work.

<details>
<summary>Definition</summary>
```{r}
data <- 
  vega_data$countries() %>%
  select(-`_comment`) %>%
  filter(
    country %in% c("China", "India", "United States", "Indonesia", "Brazil"),
    year %in% c(1955, 2000)
  ) 
```
</details>

```{r}
glimpse(data)
```

#### Chart

```{r}
# Line between life expectancy in 1955 & 2000
chart_line <-
  alt$Chart()$
  mark_line(color = "#db646f")$
  encode(
    x = "life_expect:Q",
    y = "country:N",
    detail = "country:N"
  )

# Points for life expectancy in 1955 & 2000
chart_point <- 
  alt$Chart()$
  mark_point(size = 100, opacity = 1, filled = TRUE)$
  encode(
    x = "life_expect:Q",
    y = "country:N",
    color = alt$Color(
      "year:O",
      scale = alt$Scale(
        domain = list("1955", "2000"), 
        range = list("#e6959c", "#911a24")
      )
    )
  )$interactive()

# Compose charts, add data and transformations
chart <- 
  (chart_line + chart_point)$
  properties(data =  vega_data$countries$url)$
  transform_filter(
    filter = list(
      field = "country",
      oneOf = list("China", "India", "United States", "Indonesia", "Brazil")
    )
  )$
  transform_filter(
    filter = list(field = "year", oneOf = list(1955, 2000))
  )
  
chart
```

## Steam and Leaf Plot

[Altair Example](https://altair-viz.github.io/gallery/stem_and_leaf.html)

#### Data

<details>
<summary>Definition</summary>
```{r}
data <- 
  tibble(
    sample = rnorm(100, 50, 15) %>% round() %>% as.integer(),
    stem = (sample / 10L) %>% as.integer(),
    leaf = sample %% 10L
  ) %>%
  arrange(sample) %>%
  group_by(stem) %>%
  mutate(position = seq_along(leaf)) %>%
  ungroup()
```
</details>

```{r}
glimpse(data)
```

#### Chart

```{r}
chart <- 
  alt$Chart(r_to_py(data))$
  mark_text(align = "left", baseline = "middle", dx = -5)$
  encode(
    x = alt$X(
      "position:Q",
      axis = alt$Axis(title="", ticks = FALSE, labels = FALSE, grid = FALSE)
    ),
    y = alt$Y("stem:N", axis = alt$Axis(title = "", tickSize = 0)),
    text = "leaf:N"
  )$
  configure_axis(labelFontSize = 20)$
  configure_text(fontSize = 20)

chart
```

## Trail Marker

[Altair Example](https://altair-viz.github.io/gallery/trail_marker.html)

> The `trail` marker is like the `line` marker, but it allows properties of the line (such as thickness) to vary along the length of the line. This shows a simple example of the trail mark using stock prices.

#### Data

```{r}
glimpse(vega_data$stocks())
```

#### Chart

```{r}
chart <-
  alt$Chart(stocks)$
  mark_trail()$
  encode(
    x = "date:T",
    y = "price:Q",
    size = "price:Q",
    color = "symbol:N"
  )

chart
```




