---
title: "Simple Charts"
author: "Haley Jeppson"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This document is adapted from the [Simple Charts](https://altair-viz.github.io/gallery/#simple-charts) section of the Altair Example Gallery.

Our first step is to set up our environment:

```{r}
# devtools::install_github("vegawidget/altair")
library("altair")
library("tibble")

vega_data <- import_vega_data()
```

## Simple Bar Chart

[Altair example](https://altair-viz.github.io/gallery/simple_bar_chart.html)

#### Data

<details>
<summary>Definition</summary>
```{r}
dat <-  
  data_frame(
    a= c("A", "B", "C", "D", "E", "F", "G", "H", "I"),
    b =c(28, 55, 43, 91, 81, 53, 19, 87, 52)
  )
```
</details>

```{r}
glimpse(dat)
```

#### Chart

```{r}
chart <- 
  alt$Chart(r_to_py(dat))$
  mark_bar()$
  encode(
    x = "a",
    y = "b"
  )

chart
```

## Simple Heatmap

[Altair example](https://altair-viz.github.io/gallery/simple_heatmap.html)

#### Data

<details>
<summary>Definition</summary>
```{r}
# Compute x^2 + y^2 across a 2D grid
x = -5:5
y = -5:5
dat <- expand.grid(x,y)
names(dat) <- c("x", "y")
dat$z = (dat$x^2) + (dat$y^2)
```
</details>

```{r}
glimpse(dat)
```

#### Chart

```{r}
chart <- 
  alt$Chart(r_to_py(dat))$
  mark_rect()$
  encode(
    x = "x:O",
    y = "y:O",
    color = "z:Q"
  )

chart
```

## Simple Histogram

[Altair example](https://altair-viz.github.io/gallery/simple_histogram.html)

#### Data

```{r}
glimpse(vega_data$movies())
```

#### Chart

```{r}
chart <- 
  alt$Chart(r_to_py(vega_data$movies()))$
  mark_bar()$
  encode(
    alt$X("IMDB_Rating", bin = TRUE), 
    y = "count()"
  )

chart
```

## Simple Line Chart

[Altair example](https://altair-viz.github.io/gallery/simple_line_chart.html)

#### Data

<details>
<summary>Definition</summary>
```{r}
dat <- 
  data_frame(
    x = 0:100,
    sinx = sin(x / 5)
  ) 
```
</details>

```{r}
glimpse(dat)
```

#### Chart

```{r}
chart <- 
  alt$Chart(r_to_py(dat))$
  mark_line()$
  encode(
    x = "x",
    y = "sinx"
  )

chart
```

## Simple Scatter Plot

[Altair example](https://altair-viz.github.io/gallery/simple_scatter.html)

#### Data

```{r}
glimpse(vega_data$iris())
```

#### Chart

```{r}
chart <- 
  alt$Chart(r_to_py(vega_data$iris()))$
  mark_point()$
  encode(
    x = "petalWidth",
    y = "petalLength",
    color = "species"
  )$
  interactive()

chart
```

## Simple Stacked Area Chart

[Altair example](https://altair-viz.github.io/gallery/simple_stacked_area_chart.html)

#### Data

```{r}
glimpse(vega_data$unemployment_across_industries())
```

#### Chart

```{r}
chart <- 
  alt$Chart(r_to_py(vega_data$unemployment_across_industries()))$
  mark_area()$
  encode(
    x="date:T",
    y="count:Q",
    color="series:N"
  )

chart
```

## Strip Plot

[Altair example](https://altair-viz.github.io/gallery/strip_plot.html)

#### Data

```{r}
glimpse(vega_data$cars())
```

#### Chart

```{r}
chart <- 
  alt$Chart(r_to_py(vega_data$cars()))$
  mark_tick()$
  encode(
    x = "Horsepower:Q",
    y = "Cylinders:O"
  )

chart
```







