---
title: "Maps"
author: "Haley Jeppson"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This document is adapted from the [Bar Charts](https://altair-viz.github.io/gallery/index.html#bar-charts) section of the Altair Example Gallery.

Our first step is to set up our environment:

```{r}
# devtools::install_github("vegawidget/altair")
library("altair")
library("tibble")
library("purrr")
library("dplyr")

vega_data <- import_vega_data()
```

## Choropleth Map 

[Altair example](https://altair-viz.github.io/gallery/choropleth.html)

It is not straightforward to preview the topological data, or the unemployment data:

- The `vega_data$us_10m` data is in topoJSON, not a data frame
- The `vega_data$unemployment` data is in an unparsed format.

Furthermore, we are using the URLs, which means that this example will not render in the RStudio IDE, although it will render in an external browser. 

#### Data

```{r}
counties <- alt$topo_feature(vega_data$us_10m$url, "counties")
unemp_data <- vega_data$unemployment$url
```

#### Chart

```{r}
chart <- 
  alt$Chart(counties)$
  mark_geoshape()$
  encode(
    color = "rate:Q"
  )$
  transform_lookup(
    lookup = "id",
    from_ = alt$LookupData(unemp_data, "id", list("rate"))
  )$
  project(type = "albersUsa")$
  properties(width = 500, height = 300)

chart
```

## Locations of US Airports 

[Altair example](https://altair-viz.github.io/gallery/airports.html)

#### Data

```{r}
us <- vega_data$us_10m$url
```

```{r}
airports <- vega_data$airports()
glimpse(airports)
```

#### Chart

```{r}
states <- alt$topo_feature(us, feature = "states")

# US states background
background <-
  alt$Chart(states)$
  mark_geoshape(
    fill = "lightgray",
    stroke = "white"
  )$
  properties(width = 500, height = 300)$
  project("albersUsa")

# airport positions on background
points <- 
  alt$Chart(airports)$
  mark_circle()$
  encode(
    longitude = "longitude:Q",
    latitude = "latitude:Q",
    size = alt$value(10),
    color = alt$value("steelblue")
  )

chart <- (background + points)

chart
```

## London Tube Lines 

[Altair example](https://altair-viz.github.io/gallery/london_tube.html)

> This example shows the London tube lines against the background of the borough boundaries. It is based on the [Vega-Lite example](https://vega.github.io/vega-lite/examples/geo_layer_line_london.html).

#### Data

```{r}
boroughs <- alt$topo_feature(vega_data$londonBoroughs$url, "boroughs")
tubelines <- alt$topo_feature(vega_data$londonTubeLines$url, "line")

centroids <- vega_data$londonCentroids()
glimpse(centroids)
```

#### Chart

```{r}
domain_line <- 
  list("Bakerloo", "Central", "Circle", "District", "DLR", 
       "Hammersmith & City", "Jubilee", "Metropolitan", "Northern",
       "Piccadilly", "Victoria", "Waterloo & City" )

range_line <- 
  list("rgb(137,78,36)", "rgb(220,36,30)", "rgb(255,206,0)",
       "rgb(1,114,41)", "rgb(0,175,173)", "rgb(215,153,175)",
       "rgb(106,114,120)", "rgb(114,17,84)", "rgb(0,0,0)",
       "rgb(0,24,168)", "rgb(0,160,226)", "rgb(106,187,170)")

background <-
  alt$Chart(boroughs)$
  mark_geoshape(stroke = "white", strokeWidth = 2)$
  encode(
    color = alt$value("#eee")
  )$
  properties(width = 700, height = 500)

labels <- 
  alt$Chart(centroids)$
  mark_text()$
  encode(
    longitude = "cx:Q",
    latitude = "cy:Q",
    text = "bLabel:N",
    size = alt$value(8),
    opacity = alt$value(0.6)
  )$
  transform_calculate("bLabel", "indexof (datum.name,' ') > 0  ? substring(datum.name,0,indexof(datum.name, ' ')) : datum.name")

line_scale <- 
  alt$Scale(domain = domain_line, range = range_line)

lines <- 
  alt$Chart(tubelines)$
  mark_geoshape(filled = FALSE, strokeWidth = 2)$
  encode(
    color = alt$Color(
      "id:N",
      legend = alt$Legend(
      title = NULL,
      orient = "bottom-right",
      offset = 0
    )
  )
)

chart <- (background + labels + lines)

chart
```

## One Dot Per Zipcode

[Altair example](https://altair-viz.github.io/gallery/one_dot_per_zipcode.html)

If you are building a chart using a local data frame, the default is that there has to be less than 5000 observations. If you build a chart using a URL (local or remote), there is no such default.

To modify this default:

```r
alt$data_transformers$enable('default', max_rows=<your new max>)
```

#### Data

```{r}
glimpse(vega_data$zipcodes())
```

#### Chart

```{r}
chart <- 
  alt$Chart(vega_data$zipcodes$url)$
  mark_circle(size = 3)$
  encode(
    longitude = "longitude:Q",
    latitude = "latitude:Q",
    color = "digit:N"
  )$
  properties(
    projection = list(type = "albersUsa"),
    width = 650,
    height = 400
  )$
  transform_calculate("digit", "substring(datum.zip_code, 0, 1)")

chart
```

## Repeated Choropleth Map

[Altair example](https://altair-viz.github.io/gallery/choropleth_repeat.html)

#### Data

```{r}
pop_eng_hur <- vega_data$population_engineers_hurricanes()

glimpse(pop_eng_hur)
```

#### Chart

```{r}
states <- alt$topo_feature(vega_data$us_10m$url, "states")

variable_list <- list("population", "engineers", "hurricanes")

chart <- 
  alt$Chart(states)$
  mark_geoshape()$
  encode(
    color = alt$Color(alt$`repeat`("row"), type = "quantitative")
  )$
  transform_lookup(
    lookup = "id",
    from_ = alt$LookupData(pop_eng_hur, "id", variable_list)
  )$
  properties(width = 500, height = 300)$
  project(type = "albersUsa")$
  `repeat`(row = variable_list)$
  resolve_scale(color = "independent")

chart
```

## U.S. state capitals overlayed on a map of the U.S

[Altair example](https://altair-viz.github.io/gallery/us_state_capitals.html)

#### Data

```{r}
capitals <- vega_data$us_state_capitals()
glimpse(capitals)
```

#### Chart

```{r}
states <- alt$topo_feature(vega_data$us_10m$url, "states")

# US states background
background <- 
  alt$Chart(states)$
  mark_geoshape(fill = "lightgray",stroke = "white")$
  properties(title = "US State Capitols", width = 700, height = 400)$
  project("albersUsa")

# Points and text
hover <- 
  alt$selection(
    type = "single", 
    on = "mouseover", 
    nearest = TRUE,
    fields = list("lat", "lon")
  )

base <- 
  alt$Chart(capitals)$
  encode(
    longitude = "lon:Q",
    latitude = "lat:Q"
  )

text <- 
  base$
  mark_text(dy = -5, align = "right")$
  encode(
    alt$Text("city", type = "nominal"),
    # switched from reference
    opacity = alt$condition(hover, alt$value(1), alt$value(0))
  )

points <- 
  base$
  mark_point()$
  encode(
    color = alt$value("black"),
    # switched from reference
    size = alt$condition(hover, alt$value(100), alt$value(30))
  )$
  properties(selection = hover)

chart <- (background + points + text)

chart
```

## World Projections 

[Altair example](https://altair-viz.github.io/gallery/world_projections.html)

Please see the [Vega documentation](https://vega.github.io/vega-lite/docs/projection.html) for more details on the projections available. 


```{r}
countries <- alt$topo_feature(vega_data$world_110m$url, "countries")

base <-  
  alt$Chart(countries)$
  mark_geoshape(fill = "#666666",stroke = "white")$
  properties(width = 300, height = 180)

projections <- list("equirectangular", "mercator", "orthographic", "gnomonic")

# use purrr to fashion a facet_wrap
charts <- map(projections, ~base$project(.x)$properties(title = .x))

chart <- (charts[[1]] | charts[[2]]) & (charts[[3]] | charts[[4]])

chart
```

