---
title: "Maps"
author: "Haley Jeppson"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This document is adapted from the [Bar Charts](https://altair-viz.github.io/gallery/index.html#bar-charts) section of the Altair Example Gallery.

Our first step is to set up our environment:

```{r}
# devtools::install_github("vegawidget/altair")
library("altair")
library("tibble")
library("purrr")
library("dplyr")

vega_data <- import_vega_data()
```

## Choropleth Map 

[Altair example](https://altair-viz.github.io/gallery/choropleth.html)

It is not straightforward to preview the topological data, or the unemployment data:

- The `vega_data$us_10m` data is in topoJSON, not a data frame
- The `vega_data$unemployment` data is in an unparsed format.

#### Data

```{r}
counties <- alt$topo_feature(vega_data$us_10m$url, "counties")
unemp_data <- vega_data$unemployment$url
```

#### Chart

```{r}
chart <- 
  alt$Chart(counties)$
  mark_geoshape()$
  encode(
    color = "rate:Q"
  )$
  transform_lookup(
    lookup = "id",
    from_ = alt$LookupData(unemp_data, "id", list("rate"))
  )$
  project(type = "albersUsa")$
  properties(width = 500, height = 300)

chart
```

## Locations of US Airports 

[Altair example](https://altair-viz.github.io/gallery/airports.html)

#### Data

```{r}
us <- vega_data$us_10m$url
```

```{r}
airports <- vega_data$airports()
glimpse(airports)
```

#### Chart

```{r}
states <- alt$topo_feature(us, feature = "states")

# US states background
background <-
  alt$Chart(states)$
  mark_geoshape(
    fill = "lightgray",
    stroke = "white"
  )$
  properties(width = 500, height = 300)$
  project("albersUsa")

# airport positions on background
points <- 
  alt$Chart(airports)$
  transform_aggregate(
    latitude='mean(latitude)',
    longitude='mean(longitude)',
    count='count()',
    groupby=list('state')
  )$
  mark_circle()$
  encode(
    longitude = "longitude:Q",
    latitude = "latitude:Q",
    size = alt$Size('count:Q', title='Number of Airports'),
    color = alt$value("steelblue"),
    tooltip=list('state:N','count:Q')
  )$
  properties(
    title='Number of airports in US')

chart <- (background + points)

chart
```

## Repeated Choropleth Map

[Altair example](https://altair-viz.github.io/gallery/choropleth_repeat.html)

#### Data

```{r}
#glimpse(vega_data$population_engineers_hurricanes())
```

#### Chart

```{r}
states <- alt$topo_feature(vega_data$us_10m$url, "states")
source <- vega_data$population_engineers_hurricanes$url
variable_list <- list("population", "engineers", "hurricanes")

chart <- 
  alt$Chart(states)$
  mark_geoshape()$
  encode(
    color = alt$Color(alt$`repeat`("row"), type = "quantitative")
  )$
  transform_lookup(
    lookup = "id",
    from_ = alt$LookupData(source, "id", variable_list)
  )$
  properties(width = 500, height = 300)$
  project(type = "albersUsa")$
  `repeat`(row = variable_list)$
  resolve_scale(color = "independent")

chart
```


## World Projections 

[Altair example](https://altair-viz.github.io/gallery/world_projections.html)

Please see the [Vega documentation](https://vega.github.io/vega-lite/docs/projection.html) for more details on the projections available. 


```{r}
countries <- alt$topo_feature(vega_data$world_110m$url, "countries")

base <-  
  alt$Chart(countries)$
  mark_geoshape(fill = "#666666",stroke = "white")$
  properties(width = 300, height = 180)

projections <- list("equirectangular", "mercator", "orthographic", "gnomonic")

# use purrr to fashion a facet_wrap
charts <- map(projections, ~base$project(.x)$properties(title = .x))

chart <- (charts[[1]] | charts[[2]]) & (charts[[3]] | charts[[4]])

chart
```

