---
title: "Field Guide to Python Issues"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
editor_options:
  chunk_output_type: inline
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

There are four foundations upon which this package rests: 

- the [Altair](https://altair-viz.github.io/) Python package, to build chart specifications
- the [reticulate](https://rstudio.github.io/reticulate/index.html) R package, to provide inter-operability with Python
- the [Vega-Lite](https://vega.github.io/vega-lite/) JavaScript framework, to render chart specifications in an HTML file
- the [htmlwidgets](https://www.htmlwidgets.org/) R package, to provide inter-operability with HTML and JavaScript

This article deals with the first two items; the [Field Guide to Rendering](field-guide-rendering.html) deals with the other two.

The purpose of this document is to try to collect in one place, in a semi-organized fashion, all the fiddly-bits we have found dealing with Python stuff. If you get a cryptic Python error, check here. If you find a workaround for something that isn't here, please let us know!

## Overview

The [Altair documentation](https://altair-viz.github.io) is the best resource for learning how to create charts. In the course of building and documenting this package, we have noted a few "gotchas" and their workarounds. If you find another, please let us know!

Here's the short version:

- Where you see a `.`, use a `$` instead.

- Altair methods return a *copy* of the object. Assignment of a Python object returns a *reference*, not a copy. 

- To get a copy a of "bare" Altair object, use a `$copy()` method.

- If you have a dataset that has variables with dots in their names, e.g. `Sepal.Width`, you have to make some accommodation when referring to such names in Altiar. As a workaround, you can use square-brackets to refer to "[Sepal.Width]".

- There is an Altair Chart method called `repeat()`, which in R is a reserved word, so it needs to be enclosed in backticks: ``$`repeat`()``. 

- Where you see an inversion operator, `~`, like `~highlight`, in Altair examples, call the method explicitly from R: ``hightlight$`__invert__`()``. Alternatively, you may be able to rearrange the code so as to avoid using the inversion. 

- Where you see a hyphen in the name of a Python object, use an underscore in R: `vega_data$sf_temps()`

- Where you see a Python list, `["foo", "bar"]`, in Altair examples, use an unnamed list in R: `list("foo", "bar")`.

- Where you see a Python dictionary, `{'a' = "foo", 'b' = "bar"}`, in Altair examples, use a named list in R: `list(a = "foo", b = "bar")`

- Where you see a `None` in Altair examples, use a `NULL` in R.

- You may see a function call with `**`, `baz(a = 1, **{'foo': 'bar'})`, in an Altair example. In R, interpolate the dictionary into the rest of the arguments, `baz(a = 1, foo = "bar")`.

## Data frames

Consider this Python example:

```python
# Python

from vega_datasets import data
cars = data.cars()

alt.Chart(cars).mark_point().encode(
    x='Horsepower',
    y='Miles_per_Gallon',
    color='Origin',
).interactive()
```

In this case, we are supplying a Data Frame to a `Chart()` method. 

```{r}
library("altair")

vega_data <- import_vega_data()
cars <- vega_data$cars()

chart <- 
  alt$Chart(cars)$
  mark_point()$
  encode(
    x = "Horsepower:Q",
    y = "Miles_per_Gallon:Q",
    color = "Origin:N"
  )
```

## Method chaining

When **reticulate** returns a Python object with a custom class, it appears in R as an S3 object that *behaves* like a reference class. This means that if you see this sort of notation in Python:

```python
# Python
foo.bar()
```

You would use this notation in R:

```r
foo$bar()
```

In essence, you wherever you see a `.` in Python, use a `$` in R. 

```{r cars}
vega_data <- import_vega_data()
cars <- vega_data$cars()

chart <- 
  alt$Chart(cars)$
  mark_point()$
  encode(
    x = "Horsepower:Q",
    y = "Miles_per_Gallon:Q",
    color = "Origin:N"
  )
```

## Altair method returns copy

In Python, Altair methods return a copy of the object. To verify this, let's use `pryr::

```{r}
library("pryr")

chart_new <- chart$mark_point()

address(chart_new) == address(chart)
```

Although this *looks* like a reference-class method, the Altair method *acts* like an S3 method.

## Python assignment returns reference

The object returned by an Altair method is a modified copy of the calling-object, much as we are accustomed-to in R. However, it is important to note that using the R assignment operator (`<-`, `=`, `->`) on a Python object returns a *reference* to the object rather than a copy. 

This becomes apparent when assigning a "bare" object:

```{r}
chart_new <- chart

address(chart_new) == address(chart)
```

To return a copy of the object, use a copy method.

```{r}
chart_new <- chart$copy()

address(chart_new) == address(chart)
```

## Dots in variable names

In Python, dots can refer to a nested structure within a Data Frame variable. Vega-Lite supports such nesting, so it assumes that a dot in a variable-name will refer to a nested variable. 

This means that we can run into trouble using R's `freeny` dataset:

```{r eval = FALSE}
# does not render properly

chart_freeny_r <-
  alt$Chart(freeny)$
  encode(
    x = alt$X("income.level:Q", zero = FALSE),
    y = alt$Y("market.potential:Q", zero = FALSE)
  )$
  mark_point()

chart_freeny_r
```

The problem here is that there are variables whose names have dots in them, e.g. `income.level`. One workaround is to use square brackets when referring to such variable names; another is to use backslashes, `\\`:

```{r}
chart_freeny_r <-
  alt$Chart(freeny)$
  encode(
    x = alt$X("[income.level]:Q", scale = alt$Scale(zero = FALSE)),
    y = alt$Y("market\\.potential:Q", scale = alt$Scale(zero = FALSE))
  )$
  mark_point()

chart_freeny_r
```

As you can see, this has the side-effect of showing the brackets and slashes in the scale labels.

To fix the fix, you can set the title for each axis:

```{r}
chart_freeny_r <-
  alt$Chart(freeny)$
  encode(
    x = alt$X(
      "[income.level]:Q", 
      scale = alt$Scale(zero = FALSE),
      axis = alt$Axis(title = "income.level")
    ),
    y = alt$Y(
      "market\\.potential:Q", 
      scale = alt$Scale(zero = FALSE),
      axis = alt$Axis(title = "market.potential")
    )
  )$
  mark_point()

chart_freeny_r
```

## Repeat

As shown in the [View Composistion article](view-composition.html#repeat), you can use the `repeat()` method to compose one-or-more charts such that the only thing different among them is an encoding.

However, the article notes, there is a catch: `repeat` is a reserved word in R, so we have to enclose it in backticks, e.g. ``$`repeat`()``.

```{r repeat}
chart_repeat <- 
  alt$Chart(freeny)$
  encode(
    x = alt$X(
      "[income.level]:Q",
      scale = alt$Scale(zero = FALSE),
      axis = alt$Axis(title = "income.level")      
    ),
    y = alt$Y(
      alt$`repeat`("column"), 
      type = "quantitative",
      scale = alt$Scale(zero = FALSE)
    )
  )$
  mark_point()$
  properties(
    width = 200,
    height = 200
  )$
  `repeat`(
    column = list("[market.potential]", "[price.index]")
  )

chart_repeat
```

As you can see, the `repeat` operator does not give us a way to customize the axis titles.

## Inversion: `~`

This is another case where an operator has a completely different meaning in Python than it has in R. As you know, the `~` operator is used to construct a formula. In Python, it is the [bitwise inversion operator](https://stackoverflow.com/questions/8305199/the-tilde-operator-in-python). 

You might come across this in an [Altair example](https://altair-viz.github.io/gallery/multiline_highlight.html) where the operator is used to invert a selection.

```python
# Python
highlight = alt.selection_single(on='mouseover', fields=['symbol'], nearest=True)

alt.condition(~highlight, alt.value(1), alt.value(3))
```

There are a couple of alternatives available here, the first is to invoke the `$__invert__()` operator explicitly.

```{r eval = FALSE}
# R
highlight <-
  alt$selection_single(
    on = "mouseover",
    fields = list("symbol"), 
    nearest = TRUE
  )

alt$condition(highlight$`__invert__`(), alt$value(1), alt$value(3))
```

The second alternative is to swap the order of the `if_true` and `if_false` arguments in `alt$condition()`.

```{r eval = FALSE}
# R
highlight <-
  alt$selection_point(
    on = "mouseover",
    fields = list("symbol"), 
    nearest = TRUE
  )

alt$condition(highlight, alt$value(3), alt$value(1))
```

## Hyphens in Python Names

This comes up in Vega datasets. Let's use the `$list_datasets()` method to get the names of the datasets that contain a hyphen.

```{r datasets-hyphen}
vega_data$list_datasets() %>% stringr::str_subset("-")
```

To refer to one of these datasets in R, substitute the hyphen with an underscore:

```{r}
vega_data$sf_temps() %>% head()
```

## Lists: `[]` and Dictionaries: `{}`

A Python list corresponds to an atomic vector in R; a Python dictionary corresponds to a named list in R.

```python
# Python

example_list = [1, 2, 3]
example_dictionary = {'a': 1, 'b': 2, 'c': 3}
```

In practice, we find that reticulate does the right thing if we provide an R *unnamed* list where Altair expects a list, and an R *named* list where Altair expects a dictionary.

```{r}
example_list <- list(1, 2, 3)
example_dictionary <- list(a = 1, b = 2, c = 3)
```

Consider this Altair [example](https://altair-viz.github.io/gallery/interactive_layered_crossfilter.html) that uses lists and dictionaries. These are some of the Python bits:

```python
import altair as alt
from vega_datasets import data

flights = alt.UrlData(data.flights_2k.url,
                      format={'parse': {'date': 'date'}})

brush = alt.selection_interval(encodings=['x'])
```

Here's an R translation of the complete example, which demonstrates interactive cross-filtering.

```{r}
flights <-
  alt$UrlData(
    vega_data$flights_2k$url,
    format = list(parse = list(date = "date"))
  )

brush <- alt$selection_interval(encodings = list("x"))

# Define the base chart, with the common parts of the
# background and highlights
base <- 
  alt$Chart(width = 180, height = 130)$
  mark_bar()$
  encode(
    x = alt$X(alt$`repeat`("column"))$bin(maxbins = 20),
    y = "count()"
  )

# gray background with selection
background <- base$encode(color = alt$value("#ddd"))$add_params(brush)

# blue highlights on the transformed data
highlight <-base$transform_filter(brush)

# layer the two charts & repeat
alt$
  layer(background, highlight, data = flights)$
  transform_calculate("time", "hours(datum.date)")$
  `repeat`(column = list("distance", "delay", "time"))
```

## None and `**{}`

These concepts are not related other that they are found in the same example:

```python
import altair as alt
import pandas as pd

activities = pd.DataFrame({'Activity': ['Sleeping', 'Eating', 'TV', 'Work', 'Exercise'],
                           'Time': [8, 2, 4, 8, 2]})

alt.Chart(activities).mark_bar().encode(
    alt.X('PercentOfTotal:Q', axis=alt.Axis(format='.0%')),
    y='Activity:N'
).transform_window(
    window=[alt.WindowFieldDef(op='sum', field='Time', **{'as': 'TotalTime'})],
    frame=[None, None]
).transform_calculate(
    PercentOfTotal="datum.Time / datum.TotalTime"
)
```

In this example, we have a list containing `None`, which reticulate associates with R's `NULL`. 

We also have some syntax, `**{'as': 'TotalTime'}`. This is a mechanism to [pass additional arguments](https://softwareengineering.stackexchange.com/a/131415) to a Python function, perhaps similar to `...` in R. It is passing a dictionary, so perhaps we can add the additional named argument in R:  

```{r list-dict}
library("tibble")

activities <- 
  tibble(
    Activity = c("Sleeping", "Eating", "TV", "Work", "Exercise"),
    Time = c(8, 2, 4, 8, 2)
  )

chart <- 
  alt$Chart(activities)$
  mark_bar()$
  encode(
    x = alt$X("PercentOfTotal:Q", axis = alt$Axis(format =".0%")),
    y = "Activity:N"
  )$
  transform_window(
    window = list(
      alt$WindowFieldDef(op = "sum", field = "Time", as = "TotalTime")
    ),
    frame = list(NULL, NULL)
  )$transform_calculate(
    PercentOfTotal = JS("datum.Time / datum.TotalTime")
  )

chart
```
