---
title: "Field Guide to Rendering Charts"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
editor_options:
  chunk_output_type: inline
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

There are four foundations upon which this package rests: 

- the [Altair](https://altair-viz.github.io/) Python package, to build chart specifications
- the [reticulate](https://rstudio.github.io/reticulate/index.html) R package, to provide inter-operability with Python
- the [Vega-Lite](https://vega.github.io/vega-lite/) JavaScript framework, to render chart specifications in an HTML file
- the [htmlwidgets](https://www.htmlwidgets.org/) R package, to provide inter-operability with HTML and JavaScript

This article deals with the second two items; the [Field Guide to Python Issues](field-guide-python.html) deals with the first two.

## Overview

The purpose of this document is to show how to render a chart-specification as HTML. We can use the Altair Python package to build a chart-specification; we can use the `vegawidget()` function to render it.

Here's the short version:

- Given a chart-specification `chart`, you can render it using `vegalite(chart)`, `print(chart)` or just `chart`. 

- You can examine a chart-specification using the `examine()` function, which wraps `listviewer::jsonedit()`.
  
- The `vegalite()` function has arguments `embed_options`, and `width` & `height`.

- The `embed_options` can be set using the `vega_embed()` function.

  - The principal options are `renderer`: either `"canvas"` or `"svg"`, and `actions`: a list of logical values to indicate if `export`, `source`, and `editor` links are to be included.
  
  - In your R session, you can set an `option()` called `altair.embed_options` to store your defaults for `embed_options`.
  
- The default interpretation of `width` and `height` depend on where they are set.

  - If you set the `width` and `height` in the chart-specification, this sets dimensions of the plotting area for a given  chart. The plotting area does not include axes, labels, etc., so the overall chart will be larger.
  
  - If you have a single-view chart, i.e. it is not faceted or concatenated, you can set the overall `width` and `height` using the `vegawidget()` function, overriding the chart-specification. 

  - `width` and `height` have meaning only as number of pixels.

- In R Markdown, the `vegawidget()` `width` and `height` can be set using the knitr options `vega.width` and `vega.height`, subject to the single-view constraint described above.


In this article, we will render a chart a bunch of different ways, each highlighting a different aspect of the rendering process. Let's define our base chart:

```{r base-chart}
library("altair")
library("tibble")

vega_data <- import_vega_data()

chart <- 
  alt$Chart(r_to_py(vega_data$cars()))$
  encode(
    x = "Miles_per_Gallon:Q",
    y = "Horsepower:Q",
    color = "Origin:N"    
  )$
  mark_point()
```

## Chart specification {#chart}

The simplest useful call to `vegawidget()` has only the chart specification. This is equivalent to calling `print(chart)`, or even just `chart`. 

```{r vega-chart}
## These calls are equivalent:

# print(chart)
# chart
vegawidget(chart)
```

As you can see, this renders this chart using some defaults:

- The size of the plotting area is 400 $\times$ 300 pixels (default for two quantitative axes).
- The chart is rendered using canvas.
- There are action links to export, to view source of, and to edit the chart.

You might have noticed that we did not set the size of the chart anywhere, to see where this is specified within the chart, we can use the `examine()` function: 

```{r examine-default, out.height=300}
examine(chart)
```

By expanding `config` and `view`, you will see the `height` and `width` properties. These defaults were added to the specification by Altair, on the Python side. Although we will discuss this more in the [dimensions](#dimensions) section, we note that these refer to the dimensions of the **plotting area** rather than the dimensions of the entire chart, including axes, labels, etc.

## Embedding options {#embedding}

The `embed` argument expects a `vega_embed` object, which is made by the `vega_embed()` function. 

If `embed` is `NULL`, `vegawidget()` sets it to `getOption("altair.embed_options")` which you can specify in your `.Rprofile`, as described in greater detail below. If `embed` is still `NULL`, this will result in the [vega-embed defaults](https://github.com/vega/vega-embed#options) being used: canvas rendering and action-links added below the chart. 

The principal arguments to `vega_embed()` deal with rendering and action-links: `renderer` and `actions`. 

To exclude all the action-links, set `actions = FALSE`:

```{r actions-false}
vegawidget(chart, embed = vega_embed(actions = FALSE))
```

Here's another example, rendering using SVG and using the `only_actions()` function to include only the "export" action-link:

```{r actions-custom}
vegawidget(
  chart, 
  embed = vega_embed(renderer = "svg", actions = only_actions(export = TRUE))
)
```

As with chart-specifications, you can `examine()` embedding-specifications.

If there is a default that you would like to use, you can set an option in your `.Rprofile` file. For example:

```r
options(
  repos = c(CRAN = "https://cloud.r-project.org/"),
  ...
  altair.embed_options =
    altair::vega_embed(actions = altair::only_actions(export = TRUE))
)
```

## Chart dimensions {#dimensions}

Let's look at a couple of charts where we specify the size. First, we take our `chart` and add `width` and `height` to the chart specification, which will dictate the size of the *plotting rectangle*:

```{r}
chart_size <- chart$properties(width = 300, height = 300)

chart_size
```

Contrast this with setting the dimensions using `vegawidget()`, which will dictate the size of the *entire chart*:

```{r}
vegawidget(chart, width = 300, height = 300)
```

In a chart specification, the default interpretation of `width` and `height` refers to the dimensions of the plotting rectangle. Using the chart specification, there is also a way to to [autosize](https://vega.github.io/vega-lite/docs/size.html#autosize) charts so that the dimensions refer to the entire chart, including axes, labels, etc. This is how `vegawidget()` uses its `width` and `height` arguments to specify the dimensions of the entire chart. 

In the `vegawidget()` function, the default value of `width` and `height` is `NULL`. For rendering, this means:

- The dimensions in the chart specification are respected.
- The `style` attribute of the enclosing `<div/>` is removed, allowing the chart to dictate the size of its enclosure.

To specify the overall dimensions of a chart, you can use the `width` and `height` arguments of the `vegawidget()` function. For rendering, this means:

- The non-null dimensions overwrite the values in the chart specification.
- The chart specification is modified to [autosize](https://vega.github.io/vega-lite/docs/size.html#autosize) the chart, so the entire chart conforms to the new specifications.
- The `style` attribute of the enclosing `<div/>` is removed, allowing the chart to dictate the size of its enclosure. However, because you have specified the size of the chart, you have also specified the size of its enclosure.

There are two important caveats:

1. You [cannot specify "autosize"](https://vega.github.io/vega-lite/docs/size.html#limitations), i.e. you cannot specify the overall dimensions, for any faceted, repeated, or concatenated chart.

2. The overall chart size does not take into account action-links, which generally add a height of 15-20 pixels. If you suppress the links, this is not a concern. 

This philosophy is driven by the idea that a Vega-Lite chart likes to determine its own size, and tends to resist attempts to impose size restrictions from the outside.

## R Markdown options

In R Markdown, you can set the `width` and `height` arguments of `vegawidget()` using the knitr options `vega.width` and `vega.height`. Of course, you are still subject to the "autosize" caveats outlined above.

    ```{r vega.width=300, vega.height=300}
    chart
    ```
    
```{r vega-options, echo = FALSE, vega.width=300, vega.height=300}
chart
```

We are not using the conventional knitr options, `out.width` and `out.height` because `out.width` can be set without *you* setting it explicitly, which would cause a hard-to-diagnose problem.
