---
title: "First Example"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Creating a chart

Let's make a chart, then we'll look at the bits-and-pieces we used to put it together.

```{r}
library("altair")

vega_data <- import_vega_data()

chart <- 
  alt$Chart(r_to_py(vega_data$cars()))$
  mark_point()$
  encode(
    x = "Horsepower:Q",
    y = "Miles_per_Gallon:Q",
    color = "Origin:N"
  )

vegawidget(chart)
```

### Installation

The first part of the code block is a call to load the package: 

```r
library("altair")
```

This assumes that your computer has a Python installation and an Altair installation. Please see the [Field Guide to Python](https://altair-viz.github.io/getting_started/field-guide-python.html) for more information on how to get up-and-running.

The next bit is to import the data object from the [Vega datasets](https://github.com/altair-viz/vega_datasets) Python package:

```r
vega_data <- import_vega_data()
```

### Chart object

The next part of the code block creates a chart object:

```{r create-chart}
chart <- alt$Chart(r_to_py(vega_data$cars()))
```

There's a few things going on here. The first is that the altair (R) package exposes the Altair (Python) library using the variable `alt`. In the [Altair documentation](https://altair-viz.github.io/getting_started/overview.html), the convention is to expose the Altair library through the varaiable `alt`:

```{python}
# Python
import altair as alt
```

Accordingly, we expose the "same" variable, by default, as a part of the package-loading process. 

The next step is to create the chart itself. In Python, we would use language like this:

```python
# Python
chart = alt.Chart(...)
```

Here's where [reticulate](https://rstudio.github.io/reticulate/) does its job. It exposes Python objects as R S3 objects that behave like [Reference Class](http://adv-r.had.co.nz/OO-essentials.html#rc) objects. In practical terms, wherever you see a `.` in the Altair Python documentation, use a `$` in your R code:

```r
chart <- alt$Chart(...)
```

The argument to the `Chart()` function also merits some explanation. In Python, the function expect its data argument to be a Pandas Dataframe. The reticulate package offers us a function `r_to_py()`, which we re-export in this package. This function and its complement `py_to_r()` manage conversion back-and-forth between some common data-types "shared" by R and Python. The [reticulate documentation](https://rstudio.github.io/reticulate/index.html#type-conversions) has more information.

In our case, we are converting dataframes. Although the reticulate documentation suggests (or at least I interpreted) that the conversion would be made automatically, I found it necessary to call `r_to_py()` explicitly.

The `data` argument to a chart function need not be a data frame; it can be a *reference* to a data frame like a URL, in which case you will not need to wrap in `r_to_py()`:

```{r url}
vega_data$cars$url

chart <- alt$Chart(vega_data$cars$url)
```

## Adding a mark

```{r add-mark}
chart <- 
  alt$Chart(r_to_py(vega_data$cars()))$
  mark_point()
```

In Vega, "mark" is a similar concept to "geom" in ggplot2. In this case, we are saying we want to represent our data using points.

In Python, methods can be "chained" using the `.` operator (`$` in our case), a little bit like how we use the `%>%` operator in R. To specify that we want points, we append `$mark_point()` to our Chart object. 

Note that we can extend the `$` operation across lines, giving us the illusion of piping. In the future, it may be interesting to wrap operators like `$foo` using functions that could be piped, `alt_foo(chart, ...)`.

## Adding encoding

```{r add-encoding}
chart <- 
  alt$Chart(r_to_py(vega_data$cars()))$
  mark_point()$
  encode(
    x = "Horsepower:Q",
    y = "Miles_per_Gallon:Q",
    color = "Origin:N"
  )
```

In Vega, "encoding" plays a similar role to "aesthetics" in ggplot2. We are mapping variables in the data to scales in the plot.

What we see here is, in fact, a shorthand. As explained in the [Altair documentation](https://altair-viz.github.io/user_guide/encoding.html#encoding-shorthands), there's a longer version:

```{r add-encoding-long}
chart <- 
  alt$Chart(r_to_py(vega_data$cars()))$
  mark_point()$
  encode(
    x = alt$X("Horsepower", type = "quantitative"),
    y = alt$Y("Miles_per_Gallon", type = "quantitative"),
    color = alt$Color("Origin", type = "nominal") 
  )
```

Altair recognizes [four types of data](https://altair-viz.github.io/user_guide/encoding.html#data-types), `"quantitative"`, `"nominal"`, `"ordinal"`, and `"temporal"`. 

## Displaying and Examining

Now that we have specified our chart, it remains to display it. This package provides a function `vegawidget()` that takes a chart object then renders and embeds it as an [htmlwidget](https://www.htmlwidgets.org).

```{r display}
vegawidget(chart)
```

You can examine the chart-specification by using the `examine()` function, which wraps `listviewer::jsonedit()`. To use this function, you will need to install the **listviewer** package from CRAN.

```{r examine, out.height=300}
examine(chart)
```
