---
title: "Interactive Examples"
author: "Ian Lyttle"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This document is based on the [linked-brush scatter-plot example](https://altair-viz.github.io/user_guide/selections.html#example-linked-brush-scatter-plot) found in the Altair documentation.

A goal of Vega-Lite is to implement a declaritive grammmar not only of visualization, but also of interaction. The core concept of this interactive grammar is the *selection* object. Selection mechanisms are implemented in Vega-Lite and are provided by Altair.

Let's start again with our basic plot from the [first example](first-example.html):

```{r}
library("altair")

plot_static <- 
  alt$Chart(
    r_to_py(mtcars)
  )$encode(
    x = "mpg:Q",
    y = "hp:Q",
    color = "cyl:N"
  )$mark_point(
  )

vegalite(plot_static)
```

Nothing new, so far. Our first step towards interactivity is to create a selection mechanism. This is an *interval* selecton, other types of selections are detailed in the [API reference](https://altair-viz.github.io/user_guide/API.html#altair.selection).

```{r}
brush <- alt$selection_interval()
```

We can create a new chart where the brush is bound using the selection property:

```{r}
plot_brush <- 
  plot_static$properties(
    selection = brush
  )

vegalite(plot_brush)
```

You can click-and-drag to draw a selection-box, then move it around. So far, so good, but it does not really *do* much. Let's change that. 

We need to "tell" the encoding to pay attention to the selection. We could do that by respecifying the entire chart. We introduce the `condition()` method to specify a light-gray color for observations that are not selected.

```{r}
plot_selection <-
  alt$Chart(
    r_to_py(mtcars)
  )$encode(
    x = "mpg:Q",
    y = "hp:Q",
    color =  alt$condition(brush, "cyl:N", alt$value("lightgray"))
  )$mark_point(
  )$properties(
    selection = brush
  ) 

# vegalite(plot_selection, height = 400)
```

This works, but it appears there is an easier way. It appears that you can call the `encode()` method to overwrite specific encondings for an existing chart. Thus, we have a chart where you can "see" the effect of your selection.

```{r}
plot_selection <- 
  plot_brush$encode(
    color = alt$condition(brush, "cyl:N", alt$value("lightgray"))
  )

vegalite(plot_selection)
```

Not too shabby, eh!

The capabilities become more-apparent when we combine charts. Let's create charts that look at different aspects of the data. We will use `plot_selection` as our base-chart, then modify as needed.

```{r}
plot_mpg <- 
  plot_selection$properties(
    width = 250,
    height = 250
  )

plot_disp <- 
  plot_mpg$encode(
    x = "disp:Q"
  )
```

In essence, `plot_mpg` is a smaller version of `plot_selection`; `plot_disp` is derived from `plot_mpg`, showing displacement on its x-axis, rather than mpg. It remains to compose these into a compound chart and display:

```{r}
plot_compound <- (plot_mpg | plot_disp)

vegalite(plot_compound)
```

You can select observations from either side. This is just a taste of what we can do using Vega-Lite, using the Altair API.
