---
title: "Rendering Charts"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
editor_options:
  chunk_output_type: inline
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `vegalite()` function renders a chart specification into an HTML `<div/>`.

* [**Chart specification**](#chart) (`chart`)
* [**Tooltip specification**](#tooltip) (`tooltip`)
* [**Embedding specification**](#embedding) (`embed`) 
* [**Chart dimensions**](#dimensions) (`width`, `height`) 

In this article, we will render a chart a bunch of different ways, each highlighting a different aspect of the rendering process. Let's define our base chart:

```{r base-chart}
library("altair")
library("tibble")

cars <- vega_data$cars()

base_chart <- 
  alt$Chart(r_to_py(cars))$
  encode(
    x = "Miles_per_Gallon:Q",
    y = "Horsepower:Q",
    color = "Origin:N"    
  )$
  mark_point()
```


## Chart specification {#chart}

The simplest useful call to `vegalite()` has only the chart specification: 

```{r vega-chart}
vegalite(base_chart)
```

As you can see, this renders the chart using some defaults:

- The size of the plotting area is 400 $\times$ 300 pixels.
- There are no tooltips.
- The chart is rendered using canvas.
- There are action links to export, to view source of, and to edit the chart.

You might have noticed that we did not set the size of the chart anywhere, to see where this is specified within the chart, we can use the `examine()` function: 

```{r examine-default, out.height=300}
examine(base_chart)
```

By expanding `config` and `view`, you will see the `height` and `width` properties. These defaults were added to the specification by Altair, on the Python side. Although we will discuss this more in the [dimensions](#dimensions) section, we note that these refer to the dimensions of the **plotting area** rather than the dimensions of the entire chart, including axes, labels, etc.

## Tooltip specification {#tooltip}

The `tooltip` argument expects a `vega_tooltip` object, which is made by the `vega_tooltip()` family of functions. The default is an empty call to `vega_tooltip()`, which results in no tooltips being displayed.

To display a tooltip that contains only the encoding variables, use the `vega_tooltip_encoding()` function:

```{r tooltip-encoding}
vegalite(base_chart, tooltip = vega_tooltip_encoding())
```

To display a tooltip that contains all variables in the data, use the `vega_tooltip_all()` function:

```{r tooltip-all}
vegalite(base_chart, tooltip = vega_tooltip_all())
```

Finally, to build a custom tooltip, use the `vega_tooltip()` function with `add_field()`:

```{r}
tooltip_custom <- 
  vega_tooltip() %>%
  add_field(field = "Name") %>%
  add_field(field = "Origin") %>%
  add_field(field = "Horsepower") %>%
  add_field(field = "Miles_per_Gallon", title = "Miles per Gallon")

vegalite(base_chart, tooltip = tooltip_custom)
```

You can examine a tooltip specification using the `examine()` funciton.

For more information on the options, see the [Tooltip article](tooltip.html) and the Vega Tooltip documentation. One thing to keep in mind is that you get exactly one tooltip-specification in a rendering, regardless if the chart is a single chart or a repeated chart.

## Embedding specificaiton {#embedding}

The `embed` argument expects a `vega_embed` object, which is made by the `vega_embed()` function. 

If `embed` is `NULL`, `vegalite()` sets it to `getOption("altair.embed_options")` which you can specify in your `.Rprofile`, as described in greater detail below. If `embed` is still `NULL`, this will result in the [vega-embed defaults](https://github.com/vega/vega-embed#options) being used: canvas rendering and action-links added below the chart. 

The principal arguments to `vega_embed()` deal with rendering and action-links: `renderer` and `actions`. 

To exclude all the action-links, set `actions = FALSE`:

```{r actions-false}
vegalite(base_chart, embed = vega_embed(actions = FALSE))
```

Here's another example, rendering using SVG and using the `only_actions()` function to include only the "export" action-link:

```{r actions-custom}
vegalite(
  base_chart, 
  embed = vega_embed(renderer = "svg", actions = only_actions(export = TRUE))
)
```

As with chart specifications and tooltip specifications, you can `examine()` embedding specifications.

If there is a default that you would like to use, you can set an option in your `.Rprofile` file. For example:

```r
options(
  repos = c(CRAN = "https://cloud.r-project.org/"),
  ...
  altair.embed_options =
    altair::vega_embed(actions = altair::only_actions(export = TRUE))
)
```

## Chart dimensions {#dimensions}

Let's look at a couple of charts where we specify the size. First, we take our `base_chart` and add `width` and `height` to the chart specification, which will dictate the size of the *plotting rectangle*:

```{r}
base_chart_size <- base_chart$properties(width = 300, height = 300)

vegalite(base_chart_size)
```

Contrast this with setting the dimensions using `vegalite()`, which will dictate the size of the *entire chart*:

```{r}
vegalite(base_chart_size, width = 300, height = 300)
```

In a chart specification, the default interpretation of `width` and `height` refers to the dimensions of the plotting rectangle. Using the chart specification, there is also a way to to [autosize](https://vega.github.io/vega-lite/docs/size.html#autosize) charts so that the dimensions refer to the entire chart, including axes, labels, etc. This is how `vegalite()` uses its `width` and `height` arguments to specify the dimensions of the entire chart. 

In the `vegalite()` function, the default value of `width` and `height` is `NULL`. For rendering, this means:

- The dimensions in the chart specification are respected.
- The `style` attribute of the enclosing `<div/>` is removed, allowing the chart to dictate the size of its enclosure.

To specify the overall dimensions of a chart, you can use the `width` and `height` arguments of the `vegalite()` function. For rendering, this means:

- The non-null dimensions overwrite the values in the chart specification.
- The chart specification is modified to [autosize](https://vega.github.io/vega-lite/docs/size.html#autosize) the chart, so the entire chart conforms to the new specifications.
- The `style` attribute of the enclosing `<div/>` is removed, allowing the chart to dictate the size of its enclosure. However, because you have specified the size of the chart, you have also specified the size of its enclosure.

There are two important caveats:

1. You [cannot specify "autosize"](https://vega.github.io/vega-lite/docs/size.html#limitations), thus you cannot specify the overall dimensions for any faceted, repeated, or concantenated chart.

2. The overall chart size does not take into account action-links, which generally add a height of 15-20 pixels. If you supress the links, this is not a concern. 

This philosophy is driven by the idea that a Vega-Lite chart likes to determine its own size, and tends to resist attempts to impose size restrictions from the outside.
