---
title: "Tooltips"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
editor_options:
  chunk_output_type: inline
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In Vega-Lite, a tooltip is an encoding to which variables are attached.

The Altair Python package offers us a shortcut to specify the tooltip encoding, by supplying the names of the variables we wish to appear in the tooltip. Altair and Vega-Lite apply some default behaviors.

```{r base-chart}
library("altair")

vega_data <- import_vega_data()

chart <- 
  alt$Chart(vega_data$cars())$
  mark_point()$
  encode(
    x = "Horsepower:Q",
    y = "Miles_per_Gallon:Q",
    color = "Origin:N",
    tooltip = c("Name", "Horsepower", "Miles_per_Gallon", "Origin")
  )

chart
```

You can use the `alt$Tooltip()` method to achieve the same outcome:

```{r long}
chart_new <- 
  chart$
  encode(
    tooltip = 
      list(
        alt$Tooltip(field = "Name", type = "nominal"),
        alt$Tooltip(field = "Horsepower", type = "quantitative"),
        alt$Tooltip(field = "Miles_per_Gallon", type = "quantitative"),
        alt$Tooltip(field = "Origin", type = "nominal")
      )
  )

chart_new
```

Although it is more verbose than the shorthand form, it does allow you to be more expressive:

```{r customize}
chart_new <- 
  chart$
  encode(
    tooltip = 
      list(
        alt$Tooltip(field = "Name", type = "nominal"),
        alt$Tooltip(
          field = "Year",
          title = "Year",
          type = "temporal", 
          timeUnit = "year"
        ),
        alt$Tooltip(field = "Horsepower", type = "quantitative"),
        alt$Tooltip(
          field = "Miles_per_Gallon", 
          format = ".2f",
          title = "Gas Mileage (mpg)",
          type = "quantitative"
        ),
        alt$Tooltip(field = "Origin", type = "nominal")
      )
  )

chart_new
```

If there is no aggregation or time-unit specified, Vega-Lite will use the variable-name as the title for the field. Otherwise, Vega-Lite will modify the variable-name to indicate the time-unit or the aggregation. You can use the `title` option to override the default that Vega-Lite will propose.

Also, you can specify the [format](https://vega.github.io/vega-lite/docs/format.html) of the values using the `format` option. For quantitative variables, [d3-format](https://github.com/d3/d3-format) is used; for temporal variables, [d3-time-format](https://github.com/d3/d3-time-format) is used.

## Further reading

- [Vega-Lite tooltips](https://vega.github.io/vega-lite/docs/tooltip.html)

- [Vega-Lite format](https://vega.github.io/vega-lite/docs/format.html)

- [D3 format](https://github.com/d3/d3-format)

- [D3 time format](https://github.com/d3/d3-time-format)

