---
title: "Tooltips"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
editor_options: 
  chunk_output_type: console
---

Tooltips are a [hot topic](https://github.com/altair-viz/altair/issues/240) within the Altair, Vega-Lite, and Vega community. They way tooltips are specified and rendered today may change tommorow. The purpose of this article is to focus on how things work today.

The most important thing to keep in mind is that the specification of the tooltip is *seperate* from the specification of the chart; they come together in the rendering process. Accordingly, a tooltip specification is associated with an entire chart. I imagine that this applies also to compound charts: you can provide exactly one tooltip specification.

## Show no tooltips

Let's start out with a basic chart. We have two steps here: the first step is to create a chart object using Altair methods, using `alt$Chart()`, and so on; the second step is to render the chart using `vegalite()`.

```{r basic-chart}
library("altair")

plot_basic <-
  alt$Chart(
    r_to_py(mtcars)
  )$encode(
    x = "mpg:Q",
    y = "hp:Q",
    color = "cyl:N"
  )$mark_point()

vegalite(plot_basic)
```

You will notice that the rendering does not produce tooltips. What you may not have noticed is that the `vegalite()` function has a `tooltip` argument; its default is an empty call to a helper function, `vegalite_tooltip()`. An empty call to this function tells the renderer not to produce tooltips.

## Show encoding variables

The `vegalite_tooltip()` function has a couple of friends, one of which is `vega_tooltip_encoding()`. This function is used to tell the renderer to include tooltip information for only those variables that are encoded in your plot. Here's an example:

```{r toolip-basic}
vegalite(plot_basic, tooltip = vega_tooltip_encoding())
```

## Show all variables

The other friend to `vega_tooltip()` is `vega_tooltip_all()`, which tells the renderer to include tooltip information for *all* the variables in the data - not only those encoded in the chart: 

```{r tooltip-all}
vegalite(plot_basic, tooltip = vega_tooltip_all())
```

## Customize

These first cases should cover a lot of ground. If you want to customize the tooltips, you would start with an "empty" tooltip specification, using `vega_tooltip()`. For each variable that you wish to include in the tooltip, you would add a field, using the function `add_field()`. Following [tidyverse custom](http://tidyverse.tidyverse.org/articles/manifesto.html#compose-simple-functions-with-the-pipe), you can compose a tooltip specification using the pipe operator, `%>%`:

```{r tooltip-composition}
tooltip_custom <- 
  vega_tooltip() %>%
  add_field(field = "mpg", title = "MPG") %>%
  add_field(field = "hp", title = "HP")

vegalite(plot_basic, tooltip = tooltip_custom)
```


### More options

For more options, check out the documentation for [Vega Tooltips](https://github.com/vega/vega-tooltip/blob/master/docs/customizing_your_tooltip.md). In keeping with the philosophy of this package, you should be able to map very easily between the options described in the Vega-tooltips and the arguments to `vega_tooltip()` and `add_field()`.

### JavaScript

Some of the tooltip options expect JavaScript functions. For example, the `vega_tooltip()` function has an argument, `onAppear` that asks for a JavaScript function to be called when a tooltip first appears. Consider this customization:

```{r}
tooltip_callback <-
  vega_tooltip_encoding(
    onAppear = JS("function(title,value){console.log('appear')}")
  )

vegalite(plot_basic, tooltip = tooltip_callback)
```

If you open this in your favorite broswer, open its JavaScript console, then cause a tooltip to appear in this plot, you should see the word "appear" appear in the console. 


### Helper functions

It may be useful to examine a tooltip specification, perhaps for debugging. The `examine()` function is provided to help you do this. It is a thin wrapper around the `listviewer::jsonedit()` function.

```{r tooltip-examine, out.height=300}
examine(tooltip_callback)
```

