---
title: "Vega Datasets"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
editor_options:
  chunk_output_type: inline
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The purpose of this document demonstrate some ways to use the data provided by the [Vega datasets](https://github.com/altair-viz/vega_datasets) Python package.

In the [Altair documentation](https://altair-viz.github.io/getting_started/installation.html), you will see this code used often:

```{python py-vega-datasets}
from vega_datasets import data

cars = data.cars()
```

The convention in the Altair documentation is to use the name `data` to refer to the `data` object in the `vega_datasets` package. This package offers a similar convention:

```{r r-vega-data}
library("altair")

vega_data <- import_vega_data()

cars <- vega_data$cars()
```

Our `vega_data` object has method to list oll its datasets:

```{r list-datasets}
vega_data$list_datasets() %>% head()
```

Each dataset has accessed using a function with a name being an element returned from `list_datasets()`. 

```{r anscome}
library("tibble")

vega_data$anscombe() %>% as_tibble()
```

It is useful to keep in mind that **reticulate** changes the names of the datasets, and presumably, Python objects in general. Where you see a `-` in a name of a Python object, a `_` will be used in the name of the reticulated object in R. For example, in Python: `data.sf-temps()`; in R:

```{r r-sf-temps}
vega_data$sf_temps() %>% as_tibble()
```

Each dataset has some metadata associated:

```{r anscome-description}
vega_data$anscombe$description
```

```{r anscome-is_local}
vega_data$anscombe$is_local
```

```{r}
vega_data$anscombe$references
```
