---
title: Twiddling Around
output: html_document
---

```{r}
library("altair")
library("reticulate")
library("dplyr")
library("magrittr")

knitr::opts_chunk$set(comment = "")
```

Here's how we could make a chart-spec using Python:

```{python eval = FALSE}
import altair as alt
from vega_datasets import data

iris = data.iris()

p2 = alt.Chart(iris).mark_point().encode(
    x='petalLength',
    y='petalWidth',
    color='species'
)
```

Here's an equivalent way using R.

Note that we seem to have to use `r_to_py()` - so we might consider exporting that function from __reticulate__. 

```{r}
iris_new <- 
  iris %>%
  rename(
    sepal_length = Sepal.Length,
    sepal_width = Sepal.Width,
    petal_length = Petal.Length,
    petal_width = Petal.Width,
    species = Species
  )
  
p <- altair$Chart(r_to_py(iris_new))$mark_point()$encode(
    x = "petal_length",
    y = "petal_width",
    color = "species"
)
```

## htmlwidget

To be an htmlwidget, we need three things:

1. A collection of library files, with a manifest.
2. A function to bind the widget to R.
3. Some to bind the widget to Javascript.

Let's start with what we will need for files. We have an Rmd that downloads the files and puts them into place. 

We now need a function that binds the widget to R. This function is called `vegalite()`, and it takes an `altair$Chart()` object as its main argument.

Finally, we need a Javascript-binding bit of code. I only partly understand it, but it works.

If I've done this right, then this should work:

```{r}
vegalite(p)
```

## And now for something completely different

Let's follow this [Altair tutorial on selection interactions](https://altair-viz.github.io/user_guide/selections.html).

```{r}
plot <- altair$Chart(r_to_py(mtcars))$mark_point()$encode(
    x = 'mpg:Q',
    y = 'hp:Q',
    color = 'cyl:N'
)

vegalite(plot)
```

Now, let's create a brush and bind it to the plot.

```{r}
brush <- altair$selection_interval()  # selection of type "interval"

plot <- plot$properties(selection = brush)

vegalite(plot)
```

Let's make the brush do something interesting:

```{r}
plot <- 
  altair$Chart(r_to_py(mtcars))$mark_point()$encode(
    x = 'mpg:Q',
    y = 'hp:Q',
    color = altair$condition(brush, 'cyl:N', altair$value('lightgray'))
  )$properties(selection = brush)

vegalite(plot)
```

For our next trick, link a couple of plots together.

```{r}
# also twiddling with some code-styling
plot_basic <-
  altair$Chart(
    r_to_py(mtcars)
  )$mark_point(
  )$encode(
    y = 'hp:Q',
    color = altair$condition(brush, 'cyl:N', altair$value('lightgray'))
  )$properties(
    width = 250,
    height = 250,
    selection = brush
  )

plot_left <- 
  plot_basic$encode(
    x = 'disp:Q'
  )

plot_right <- 
  plot_basic$encode(
    x = 'mpg:Q'
  )

plot_compound <- altair$hconcat(plot_left, plot_right)

vegalite(plot_compound)
```

## Your turn

Have a look at the [Altair documentation](https://altair-viz.github.io/) to see what is possible.
