---
title: "Maps"
author: "Haley Jeppson"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This document is adapted from the [Bar Charts](https://altair-viz.github.io/gallery/index.html#bar-charts) section of the Altair Example Gallery.

Our first step is to set up our environment:

```{r}
# devtools::install_github("vegawidget/altair")
library("altair")
library("tibble")
library("purrr")

vega_data <- import_vega_data()
```

## Choropleth Map (not working)

Something is not being specified correctly in the line `alt$topo_feature(r_to_py(vega_data$us_10m()), "counties")`

#### Data

```{r}
us_10m <- vega_data$us_10m()
glimpse(us_10m)
```

```{r}
unemp_data <- vega_data$unemployment()
glimpse(unemp_data)
```

#### Chart

```{r, eval = FALSE}
counties <- alt$topo_feature(r_to_py(us_10m, "counties"))

chart <- 
  alt$Chart(counties)$
  mark_geoshape()$
  encode(
    color = "rate:Q"
  )$
  transform_lookup(
    lookup = "id",
    from_ = alt$LookupData(r_to_py(unemp_data), "id", list("rate"))
  )$
  project(type = "albersUsa")$
  properties(width = 500, height = 300)

chart
```

## Locations of US Airports (not working)

#### Data

```{r}
us <- vega_data$us_10m()
glimpse(us)
```

```{r}
airports <- vega_data$airports()
glimpse(airports)
```

#### Chart

Something is not being specified correctly in the line`alt$topo_feature(us, feature="states")`

```{r, eval = FALSE}
states <- alt$topo_feature(r_to_py(us), feature = "states")

# US states background
background <-
  alt$Chart(r_to_py(states))$
  mark_geoshape(
    fill = "lightgray",
    stroke = "white"
  )$
  properties(width = 500, height = 300)$
  project("albersUsa")

# airport positions on background
points <- 
  alt$Chart(r_to_py(airports))$
  mark_circle()$
  encode(
    longitude = "longitude:Q",
    latitude = "latitude:Q",
    size = alt$value(10),
    color = alt$value("steelblue")
  )

chart <- (background + points)

chart
```

#### London Tube Lines (not working)

Need to figure out how to specify the `topo_feature`

#### Data

We don't have a good way, yet, to glimpse the data used in `alt$topo_feature()`.

```{r}
centroids <- vega_data$londonCentroids()
glimpse()
```

#### Chart

```{r, eval = FALSE}
boroughs <- alt$topo_feature(vega_data$londonBoroughs$url, "boroughs")
tubelines <- alt$topo_feature(vega_data$londonTubeLines$url, "line")
centroids <- r_to_py(centroids)

domain_line <- 
  list("Bakerloo", "Central", "Circle", "District", "DLR", 
       "Hammersmith & City", "Jubilee", "Metropolitan", "Northern",
       "Piccadilly", "Victoria", "Waterloo & City" )

range_line <- 
  list("rgb(137,78,36)", "rgb(220,36,30)", "rgb(255,206,0)",
       "rgb(1,114,41)", "rgb(0,175,173)", "rgb(215,153,175)",
       "rgb(106,114,120)", "rgb(114,17,84)", "rgb(0,0,0)",
       "rgb(0,24,168)", "rgb(0,160,226)", "rgb(106,187,170)")

background <-
  alt$Chart(boroughs)$
  mark_geoshape(stroke = "white", strokeWidth = 2)$
  encode(
    color = alt.value("#eee")
  )$
  properties(width = 700, height = 500)

labels <- 
  alt$Chart(centroids)$
  mark_text()$
  encode(
    longitude = "cx:Q",
    latitude = "cy:Q",
    text = "bLabel:N",
    size = alt$value(8),
    opacity = alt$value(0.6)
  )$
  transform_calculate("bLabel", "indexof (datum.name,' ') > 0  ? substring(datum.name,0,indexof(datum.name, ' ')) : datum.name")

line_scale <- 
  alt$Scale(domain = domain_line, range = range_line)

lines <- 
  alt$Chart(tubelines)$
  mark_geoshape(filled = FALSE, strokeWidth = 2)$
  encode(
    color = alt$Color(
      "id:N",
      legend = alt$Legend(
      title = NULL,
      orient = "bottom-right",
      offset = 0
    )
  )
)

chart <- (background + labels + lines)

chart
```

#### One Dot Per Zipcode

Here is the [issue](https://github.com/altair-viz/altair/issues/611) where there has to be less than 5000 observations


#### Data

```{r}
zips <- vega_data$zipcodes() %>% sample_n(5000)
glimpse(zips)
```

#### Chart

```{r}
chart <- 
  alt$Chart(r_to_py(zips))$
  mark_circle(size = 3)$
  encode(
    longitude = "longitude:Q",
    latitude = "latitude:Q",
    color = "digit:N"
  )$
  properties(
    projection = list(type = "albersUsa"),
    width = 650,
    height = 400
  )$
  transform_calculate("digit", "substring(datum.zip_code, 0, 1)")

chart
```

#### Repeated Choropleth Map (not working)

#### Data

```{r}
pop_eng_hur <- vega_data$population_engineers_hurricanes()

glimpse(pop_eng_hur)
```

#### Chart

```{r, eval=FALSE}
states <- alt$topo_feature(r_to_py(vega_data$us_10m$url, "states")

pop_eng_hur <- r_to_py(pop_eng_hur)

variable_list <- list("population", "engineers", "hurricanes")

chart <- 
  alt$Chart(states)$
  mark_geoshape()$
  encode(
    color = alt$Color(alt$`repeat`("row"), type = "quantitative")
  )$
  transform_lookup(
    lookup = "id",
    from_ = alt$LookupData(pop_eng_hur, "id", variable_list)
  )$
  properties(width = 500, height = 300)$
  project(type = "albersUsa")$
  `repeat`(row = variable_list)$
  resolve_scale(color = "independent")

chart
```

#### U.S. state capitals overlayed on a map of the U.S (not working)

#### Data

```{r}
capitals <- vega_data$us_state_capitals()
glimpse(capitals)
```

#### Chart

```{r, eval=FALSE}
states <- alt$topo_feature(vega_data$us_10m$url, "states")
capitals <- r_to_py(capitals)

# US states background
background <- 
  alt$Chart(states)$
  mark_geoshape(fill = "lightgray",stroke = "white")$
  properties(title = "US State Capitols", width = 700, height = 400)$
  project("albersUsa")

# Points and text
hover <- 
  alt$selection(
    type = "single", 
    on = "mouseover", 
    nearest = TRUE,
    fields = list("lat", "lon")
  )

base <- 
  alt$Chart(capitals)$
  encode(
    longitude = "lon:Q",
    latitude = "lat:Q"
  )

text <- 
  base$
  mark_text(dy = -5, align = "right")$
  encode(
    alt$Text("city", type = "nominal"),
    # switched from reference
    opacity = alt$condition(hover, alt$value(1), alt$value(0))
  )

points <- 
  base$
  mark_point()$
  encode(
    color = alt$value("black"),
    # switched from reference
    size = alt$condition(hover, alt$value(100), alt$value(30))
  )$
  properties(selection = hover)

chart <- (background + points + text)

chart
```

#### World Projections (not working)

I am unfamiliar with a lot of the python notation used in this example

```{r, eval = FALSE}

countries <- alt$topo_feature(r_to_py(vega_data$us_110m$url, "countries")

base <-  
  alt$Chart(countries)$
  mark_geoshape(fill = "#666666",stroke = "white")$
  properties(width = 300, height = 180)

projections <- list("equirectangular", "mercator", "orthographic", "gnomonic")

# this seems to be a way to make a facet_wrap
charts <- map(projections, ~base$project(.x)$properties(title = .x))

chart <- (charts[[1]] | charts[[2]]) & (charts[[3]] | charts[[4]])
```

